### 2025-04-25 Article/Article.tex
Adding test Github Actions

%%
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

%%%CONTEMPORARY%%%
\documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
%\documentclass[unnumsec,webpdf,contemporary,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,contemporary,small]{oup-authoring-template}

%%%MODERN%%%
%\documentclass[unnumsec,webpdf,modern,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,modern,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,small]{oup-authoring-template}

%%%TRADITIONAL%%%
%\documentclass[unnumsec,webpdf,traditional,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,namedate,webpdf,traditional,medium]{oup-authoring-template}
%\documentclass[namedate,webpdf,traditional,small]{oup-authoring-template}

%\onecolumn % for one column layouts

%\usepackage{showframe}
\usepackage{anyfontsize}
\usepackage{multirow}
\graphicspath{{doc/Images/}} % will allow Latex to find the images folder
\usepackage{graphicx}
\sloppy % allows line spacing stretch to justify the text
% line numbers
%\usepackage[mathlines, switch]{lineno}
%\usepackage[right]{lineno}

\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}

\begin{document}

\journaltitle{Journal Title Here}
\DOI{DOI HERE}
\copyrightyear{2022}
\pubyear{2019}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}

\firstpage{1}

%\subtitle{Subject Section}

\title[Short Article Title]{TersectBrowser+ : A single browser for in-depth introgression analysis of resequenced genome datsaets}

\author[2]{Tanya Stead}
\author[2]{David Oluwasusi}
\author[2]{Gregory Lupton}
\author[2]{Gabrielle Baumberg}
\author[1,$\ast$]{Tomasz Kurowski\ORCID{}}

\authormark{Author Name et al.}

\address[1]{\orgdiv{Centre for Soil, Agrifood and Biosciences}, \orgname{Cranfield University}, \orgaddress{\street{College Road}, \postcode{MK43 0AL}, \state{Cranfield}, \country{UK}}}
\address[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[4]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}

\corresp[$\ast$]{Corresponding author. \href{email:email-id.com}{email-id.com}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

%\editor{Associate Editor: Name}

%\abstract{
%\textbf{Motivation:} .\\
%\textbf{Results:} .\\
%\textbf{Availability:} .\\
%\textbf{Contact:} \href{name@email.com}{name@email.com}\\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Journal Name}
%online.}

\abstract{The field of cultivar genomics has evolved rapidly, with growing emphasis on mapping introgressions—segments of DNA from wild plant species that have been incorporated into cultivated lines. A variety of bioinformatic tools aim to interpret this expanding complexity of plant genomic data, predict introgressions, and identify commercially relevant regions for future breeding. Demand is increasing for tailored tools to interactively explore these complex genomic queries in real time. Created in 2022, the Web-based Tersect Browser can dynamically display introgression patterns by applying flexible set theoretical expressions to sets of sequence variant data. Our 2025 development of Tersect Browser offers multiple front-end extension capabilities so that the user can view and analyze data without leaving the page. We identify previously published introgressions with our tool, illustrating its potential for future research, and enriching plant breeders' understanding of crop cultivar genomics.}
\keywords{introgression, browser, resequenced, SNP}

% \boxedtext{
% \begin{itemize}
% \item Key boxed text here.
% \item Key boxed text here.
% \item Key boxed text here.
% \end{itemize}}

\maketitle
\section{Introduction}

\section{Expanding landscape of plant genomic studies}\label{sec2}

The field of tomato genomics has continued to evolve rapidly, with a growing emphasis on mapping introgressions — segments of DNA from wild tomato species that have been incorporated into cultivated lines. Foundational work such as that by Aflitos et al. \cite{RN7} , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.

Recent studies have expanded on this base with increasingly sophisticated tools and datasets. Notably, Liu et al. (2023) used tomato pan-genomes built from hundreds of accessions to uncover regions of structural variation and introgression hotspots, many of which affect key traits like fruit size, shape, disease resistance, and flavour. The construction of these pan-genomes enables researchers to observe variation absent in the reference genome, improving trait mapping and introgression detection.

\subsection{Background on introgression studies}\label{subsec1}

The role of introgression in water-tolerance Kubond et al. (2022), shows how such analysis will be important in identifying variant resistance to climate change.

To manage and interpret the expanding complexity of genomic data, a wide range of bioinformatic tools have been deployed. Established platforms such as SnpEff, VEP, BEDTools, BEDOPS, CrossMap, and the UCSC Genome Browser continue to support annotation, variant effect prediction, and coordinate translation. Visualization utilities such as the Introgression Browser (Aflitos et al., 2015) and the Tersect Browser (Kurowski and Mohareb, 2019) offer tailored environments for exploring introgression patterns, with Tersect’s set-theoretic framework enabling high-speed, complex genomic queries.

Meanwhile, long-read sequencing technologies (e.g., PacBio, Hi-C, and optical mapping) as applied in SL4.0 (Hosmani et al., 2019), and innovations in in situ hybridization (Shearer et al., 2014), have markedly improved the resolution of reference genomes—critical for detecting subtle and ancient introgressions.

Beyond tomato, research in soybean introgression offers valuable cross-crop insights. Studies have shown how introgressed genomic regions from wild Glycine soja into Glycine max improve traits like drought resistance, seed oil composition, and disease resilience. Tools such as SoyFGB v2.0 \footnote{\url{https://sfgb.rmbreeding.cn/index}}, SoyBase\footnote{\url{https://www.soybase.org/}}, and SoyKB\footnote{\url{http://soykb.org/}} provide functional genomics platforms that may serve as models for tomato-specific resources. Graph genome approaches and machine learning models are emerging as promising directions for future introgression prediction and trait association.

Together, these advances paint a dynamic picture of tomato genomics: one where introgression is not only a window into domestication and evolutionary history but also a powerful mechanism for ongoing improvement, including against threats such as climate change. The fusion of high-resolution genomic data, computational methods, and specialized tools enables researchers to explore tomato’s genetic landscape with increasing clarity and sophistication.  
(Introgression hybridisation is an important factor in crop improvement. Biologists need to determine introgressions and identify donor species through visualising genetic distance and phylogenetic relationships based on the whole genome variant data. However, existing software is not suitable for full exploitation of the large publicly available data sets. Tersect Browser is a Web Application that is optimised for generating such visualisations.)

\section{This is an example for first level head}\label{sec3}

\subsection{This is an example for second level head - subsection head}\label{subsec2}

\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}

\paragraph{This is an example for fourth level head - paragraph head}

\section{Conclusion}

Some Conclusions here.

%%%%%%%%%%%%%%

\begin{appendices}

\section{Section title of first appendix}\label{sec11}

\subsection{Subsection title of first appendix}\label{subsec4}

\section{Example of another appendix section}\label{sec13}%

\end{appendices}

\section{Code availability}
The technical documentation is available at XX. The source code for the project is freely accessible on GitHub at XX.

\section{Competing interests}

\section{Author contributions statement}

\section{Acknowledgments}

\bibliography{examplerefs}

\begin{thebibliography}{10}
\bibitem{bahdanau2014neural}
Dzmitry Bahdanau, Kyunghyun Cho, and Yoshua Bengio.
\newblock Neural machine translation by jointly learning to align and
  translate.
\newblock {\em arXiv preprint arXiv:1409.0473}, 2014.
\end{thebibliography}

\end{document}

### 2025-04-25 TechDoc/TechDoc.tex
Update TechDoc.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{soul}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm]{geometry}

\title{
{\includegraphics[width=3cm, height=2.5cm]{./Cran.jpg}}
\\
\includegraphics[width=7cm, height=1cm]{./TB.jpg}
\\
{Technical Documentation for TersectBrowser+}
}

\author{David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}

\begin{document}
\sloppy % allows line spacing stretch to justify the text

\maketitle

\section{Overview}
This Technical Documentation describes the particular details of updating the original TersectBrowser developed by Tomasz Kurowski into the TersectBrowser+ by integrating multiple extension features.
\section{Architecture}
The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22. It was chosen to use a 'bolt-on' approach for the extensions in order to fit the project delivery within the time requirements, as well as allowing the new extensions to use up-to-date packages. The app is dockerized to allow both versions of node to be used seamlessly in different sections.

\section{Deployment}

\subsection{System Requirements}
TersectBrowser+ can be setup and deployed from both Mac and Windows machines, with variations in the configuration required.

\section{TersectBrowser}
\section{TersectBrowser+}
\subsection{Variant Browser Extension}
\label{sec:Browser}
\subsubsection{Rationale}

\subsubsection{User Interface}

\paragraph{Feature Description}
\begin{itemize}
    \item Window is located at top of Tersect Browser (TB)- below settings bar but above TB heatmap 
    \item Jbrowse window (JB) layout - shows only sequence tracks and ruler track/linear genome scale. Zoom features/chromosome selection features have been removed from Jbrowse window. Hamburger button opens menu containing many options, notably track selector, where tracks can be added/removed. The chromosome scale for TB has a dynamic offset depending on the accession/tree view and the zoom level. This is synced between JB and TB, so that the chromosome scale lines up. 
    \begin{itemize}
        \item \hl{Layout TBC}: Set window size
    \end{itemize}
    \item Tracks:
    \begin{itemize}
        \item Reference sequence 
        \item Variant tracks for each loaded accession
        \item Gene models FeatureTrack (Ext-GM) 
    \end{itemize}
    \item Variant browser window is preloaded with default tracks. 
    \begin{itemize}
        \item \hl{Track TBC}: Currently pre-loaded with first 3 tracks from dataset. Should be changed to 1-2 meaningful accessions?
    \end{itemize}
    \item Syncing of the Browser pane and the Tersect heatmap pane, including the following:
    \begin{itemize}
        \item Zoom synced, both when the zoom plus/minus buttons are pressed, and also when scrolling the mouse to zoom. 
        \item Bin size changes synced.
        \item Interval synced using an offset of the Browser pane.
        \item Chromosome selection changes synced.
        \item Horizontal scrolling along the Browser and Tersect panes is synced.
    \end{itemize}
\end{itemize}

\subsubsection{Technical Description}

\paragraph{View State} 
The \verb +JBrowseLinearGenomeView+ component is imported from @jbrowse/react-linear-genome-view and is wrapped using React. The configurations for the assembly, assembly, and variant tracks, tracks, are imported from the respective assembly.ts and tracks.ts configuration files. The styling configurations, config, are imported from jbrowseConfig.ts. The \verb +JbrowseWrapperProps+ are imported from \verb +JbrowseInterface+.

The function \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+. If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the constant \verb +JbrowseWithAccessionName+ is returned (See below parameters). If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated. If either of them are empty, a container with the phrase "Loading…" is returned. If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+. This takes as arguments the following variables: 
\begin{itemize}
\item \verb +assembly+: the assembly track 
\item \verb +tracks+: the data tracks to display 
\item \verb +configuration+: the config variable where the theme is defined 
\item \verb +defaultSession+: an object of type LinearGenomeView, which defines the initial state. This object includes the following configurations: 
\begin{itemize}
\item \verb +bpPerPx+: the base pairs per pixel, which defines the bin size 
\item \verb +assemblyName+: the corresponding assembly for the track 
\item \verb +start+: the track start position 
\item \verb +end+: the track end position 
\item \verb +refName+: the genomic coordinates of the viewing window 
\end{itemize}
\end{itemize}

When no accession is selected in the Canvas interface, the first three tracks defined in \verb +tracks.ts+ are added to the \verb +viewState+. The dynamic left-hand offset is set by \verb +horizontalScroll()+.

\paragraph{Assembly Config}

The configurations for the assembly track, which is built upon the reference accession SL2.50, is defined in \verb +assembly.ts+ in json format. The name is set to the reference accession name and \verb +trackId+ is set to "SL2.50-ReferenceSequenceTrack". The paths to the fasta file, along with its corresponding fasta index, are specified as local server URLs provided by the backend during the Tersect Browser setup.

\paragraph{Tracks Config} 
The configurations for the \verb +VariantTrack+ and Gene Models \verb +FeatureTrack+ are defined in \verb +tracks.ts+ in json format. For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to "ITAG2.4 Gene Models".  The paths to the zipped files, along with its corresponding Tabix index files, are specified as local server URLs provided by the backend during the Tersect Browser setup. Each track is separated by a comma.

\paragraph{Styling}
The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour '\#459e00' and the \verb +boxShadow+ is set to 'none'.

\paragraph{Zoom} 
The zoom is synchronized between Tersect Browser and the JBrowse component. The \verb +zoomLevel+ observable is a component of the \verb +PlotStateService+ class. Inside \verb +tersect-browser.component.ts+, the subscription \verb +zoomSub+ listens to the \verb +zoomLevel+ observable and assigns the latest value to the component \verb +zoomLevel+. Inside \verb +tersect-browser.component.html+, the \verb +zoomLevel+ is passed to \verb +JbrowseWrapper+ as a prop and is used to define the \verb +bpPerPx+ in the \verb +viewState+.

The bin size is synchronised in the same way: the \verb +binsize+ observable is a component of the \verb +PlotStateService+ class. Inside \verb +tersect-browser.component.ts+, the subscription \verb +binSizeSub+ listens to the \verb +binsize+ observable and assigns the latest value to the component \verb +binSize+. Inside \verb +tersect-browser.component.html+, the \verb +binSize+ is passed to \verb +JbrowseWrapper+ as a prop. 
Together, \verb +bpPerPx+ is calculated with the following equation: \begin{equation} bpPerPx = ((props.location.binSize) * (100 /props.location.zoomLevel)) \end{equation}

\paragraph{Chromosome selection}
The displayed chromosome is synced in a similar fashion to the zoom and bin size. The \verb +chromosome+ observable is a component of the \verb +PlotStateService+ class. Inside \verb +tersect-browser.component.ts+, the subscription \verb +chromosomeSub+ listens to the \verb +chromosome+ observable and assigns the latest value to the object \verb +selectedChromosomeSub+. Inside \verb +tersect-browser.component.html+, the \verb +selectedChromosomeSub+ is passed to \verb +JbrowseWrapper+ as the prop chromosome. Inside \verb +JbrowseWrapper+, the chromosome name is called from the chromosome object and used to define the \verb +refName+ in the \verb +viewState+.

Additionally, the default chromosome that is pre-selected when Tersect Browser initially loads is passed to \verb +JbrowseWrapper+ and used to define the default \verb +viewState+. Inside \verb +tersect-browser.component.ts+, the variable \verb +preselectedChromosome+ is defined. On initializing, when the \verb +settings+ observer subscribes to the \verb +tersectBackendService+, the current \verb +plotState.chromosome+ is saved to the \verb +preselectedChromosome+ variable. This is then passed as the prop \verb +preselectedChromosome+ to \verb +JBrowseWrapper+ inside \verb +tersect-browser.component.html+.  Inside \verb +JbrowseWrapper+, the chromosome name is called from the \verb +preselectedChromosome+ object and used to define the \verb +refName+ in the default \verb +viewState+.

\paragraph{Interval display}
The displayed interval is synced in a similar fashion to the zoom, bin size, and chromosome selection. The \verb +interval+ observable is a component of the \verb +PlotStateService+ class. Inside \verb +tersect-browser.component.ts+, the subscription \verb +selectedIntervalSub+ listens to the \verb +interval+ observable and assigns the latest value to the array \verb +selectedInterval+. Inside \verb +tersect-browser.component.html+, the \verb +selectedInterval+ is passed to \verb +JbrowseWrapper+ as the prop \verb +selectedInterval+. Inside \verb +JbrowseWrapper+, the first element of the array is used to define the start position in the \verb +viewState+, and the second element is used to define the end position. 
Additionally, the default interval that is pre-selected when Tersect Browser initially loads is passed to \verb +JbrowseWrapper+ and used to define the default \verb +viewState+. Inside \verb +tersect-browser.component.ts+, the variable \verb +defaultInterval+ is defined. On initializing, the method \verb +generateMissingSettings+ loads the interval based on the size of the selected chromosome, which is obtained from \verb +BrowserSettings+. The interval is saved as an array to \verb +defaultInterval+ and inside \verb +tersect-browser.component.html+ it is passed as the prop \verb +defaultInterval+ to \verb +JbrowseWrapper+. Inside \verb +JbrowseWrapper+, the first element of the array is used to define the start position in the default \verb +viewState+, and the second element is used to define the end position.

\paragraph{Offset}
The dynamic offset is synced in a similar fashion to zoom, bin size, chromosome selection, and interval. The observable \verb +offsetCanvas+ is defined as a component of the \verb +PlotStateService+ class. The public variable \verb +offsetCanvasSource+ is defined as an instance of the \verb +BehaviourSubject+ class and holds all recorded values of the canvas offset. In the class constructor, \verb +offsetCanvas+ is initialised to continuously hold the latest value from \verb +offsetCanvasSource+. The canvas offset is set in the \verb +TreePlotComponent+ class, and is passed to \verb +offsetCanvasSource+ when the tree is redrawn.

Inside \verb +tersect-browser.component.ts+, the variable \verb +offsetCanvas+ is defined, and the subscription \verb +offsetCanvasSub+ listens to the  \verb +offsetCanvas+ observable and assigns the latest value to \verb +offsetCanvas+. Inside \verb +tersect-browser.component.html+, the \verb +offsetCanvas+ is passed to \verb +JbrowseWrapper+ as the prop \verb +offsetCanvas+. Inside \verb +JbrowseWrapper+, the variable defines the extent of the \verb +horizontalScroll+ using the following logic: $horizontalScroll(-(location.offsetCanvas - 4))$.

\subsubsection{Test Results}
When additional variant tracks are added to the Browser view, the size of the browser panel does not change. The user is able to scroll downwards to view more tracks in the browser.

\subsubsection{Future Improvements}
(Current) Limitations:

Only SL2.50 loaded as assembly and reference

Can only view variant tracks for 150\_VCF\_Tomato dataset - this is hard coded, and would need to be changed if we want this to be dynamic (when the user adds their own dataset) / if we as admins want to load soybean dataset!

\subsection{Gene Models Extension}
\label{sec:Models}
\subsubsection{Rationale}
\subsubsection{User Interface}
\paragraph{Feature Description}
\begin{itemize}
    \item In the Jbrowse window, the first track shows gene models based on the GFF file provided.  
    \item Gene model annotations are based on the official \hl{tomato genome annotation v2.4}: \href{https://phytozome-next.jgi.doe.gov/info/Slycopersicum_ITAG2_4#:~:text=Genome%20Overview,genome%2Cwith%2034%2C725%20gene%20models}{Slycopersicum ITAG annotation v2.4}
    \item The popup view of the same Jbrowse component has similar default: the first track is the gene models track, and the second track is the variant track for the selected accession.
\end{itemize}

\subsubsection{Technical Description}
The configuration for the Gene Models \verb+FeatureTrack+ is defined as the first entry in \verb+tracks.ts+. The paths to the sorted and compressed GFF file, along with its corresponding Tabix index, are specified as local server URLs provided by the backend during the Tersect Browser setup.

The Jbrowse \verb+viewState+ is configured in \verb+JbrowseWrapper.tsx+, as described in Extension-JBrowse. If no accession is selected, the first three tracks stored in \verb+tracks.ts+ are added to the \verb+viewState+. As the Gene Models \verb+FeatureTrack+ is the first track in \verb+tracks.ts+, it is displayed as the first track in the Jbrowse window.

For the popup window, the JBrowse \verb+viewState+ is configured when an accession is selected in \verb+JbrowseWithAccession.tsx+, as described in \nameref{sec:Browser}. If a track with a \verb+trackId+ matching the selected accession is found, the viewer displays both the first track defined in \verb+tracks.ts+ and the matching track.

\subsubsection{Test Results}
\subsubsection{Future Improvements}

\subsection{Feature Search Extension}
\subsubsection{Rationale}
The aim of this extension is to allow the user to search specific genes and identify which accession contains a high/medium/low impact variant impacting that gene.
\subsubsection{User Interface}
\paragraph{Feature Description}
\begin{itemize}
    \item Separate search bar in the top right of the tersect browser header where user can input gene name and search button 
    \item \hl{TBC: Option} for user to select based on region in addition to gene name 
    \item Popup window for user to select advanced features 
    \item Output: Bins in the canvas are highlighted red at the chromosomal position where the gene is located, and only for accessions containing a variant impacting that gene 
    \item Clear button to clear highlighted bins from canvas 
\end{itemize}
\subsubsection{Technical Description}
\paragraph{\hl{Search Bar}}
\paragraph{\hl{Popup window with advanced settings }}
\paragraph{Highlighting bins}
Bin highlighting is controlled in the \verb+bin-draw.service+ by two functions. The first function, \verb+highlightFeatureBins()+, is defined, taking as arguments a string containing accession names, the bin position along the x-axis, and the \verb+binView+. First, the y-axis bin position for accession names is calculated. Then, the \verb+binView+ is redrawn in greyscale, with the bin colour determined by the difference to the reference accession. Using the x-axis \verb+binIndex+ position and the y-axis accession bin positions, these bins are coloured red in the \verb+binView+. The modified \verb+binView+ is returned as the output. The second function, \verb+highlightBins()+, takes as arguments the start position of the interval and the bin size currently shown in the Tersect pane, a list of ordered accessions shown in the canvas, and the searched accessions. For each searched accession, accession name is reformatted to match the format displayed in the Tersect pane. Then, the bin position along the x-axis is calculated using the \verb+getBinIndexFromPosition()+ function, which takes as arguments the feature position along the chromosome, the interval start position, and the bin size. Then, \verb+highlightFeatureBins()+ is called for each accession. Lastly, the canvas is redrawn using the modified \verb+binView+ output from \verb+highlightFeatureBins()+. 
\begin{itemize}
    \item highlightFeatureBins(accessions: string[], binIndex: number, binView: DistanceBinView) - takes a string of accession names, a binIndex (corresponding to the bin position along the x-axis matching the gene position on the chromosome), and binView. The y-axis index for bins matching accession names in the accessions strings is calculated and combined with the binIndex to colour these specific bins red. The rest of the bins are coloured in greyscale, with saturation depending on binDistance (calculated from tersect on the backend 
    \begin{itemize}
        \item Bin-draw.service.ts 
        \item Called by highlightBins() in bin-draw.service.ts 
    \end{itemize}
    \item highlightBins(intervalStart, binsize, orderedAccessions, searchedAccessions) - takes selected interval start position, selected binSize, list of ordered accessions shown in the canvas, and searched accessions (passed from callback function?). Accession names in Jbrowse are in a different format to what is stored in tersect browser, so accession names are reformatted to match tersect browser bins. binIndex is calculated for the selected gene. These two are passed to highlightFeatureBins() - along with this.bins - to highlight the bins. The canvas is then redrawn using the binView calculated in highlightFeatureBins(). 
    \begin{itemize}
        \item Bin-draw.service.ts 
        \item Called by callHighlightBins() in tersect-browser.component.ts 
    \end{itemize}
    \item callHighlightBins(searchedAcessions) - takes searchedAccessions. Calls highlightBins(), passing along  selected interval start position, selected binSize, list of ordered accessions shown in the canvas, and searched accessions (passed from callback function?). 
    \begin{itemize}
        \item Tersect-browser.component.ts 
    \end{itemize}
\end{itemize}

\paragraph{Calling Highlighting Bins} \mbox{}
\\
\hl{TBC: Mechanism} of searching VCFs to identify variants

Finally, callHighlightBins() is called, which itself calls highlightBins() from the bin-draw.service, passing as arguments the current interval start position, binsize, displayed accessions in an ordered format, and the searched accessions.  
\paragraph{\hl{Clear Button }}

\subsubsection{Test Results}
\subsubsection{Future Improvements}
\subsection{Barcode Generation Extension}
\subsubsection{Rationale}
\subsubsection{User Interface}
\subsubsection{Technical Description}
\subsubsection{Test Results}
\subsubsection{Future Improvements}
\subsection{Automated Introgression Search Extension}
\subsubsection{Rationale}
\subsubsection{User Interface}
\subsubsection{Technical Description}
\subsubsection{Test Results}
\subsubsection{Future Improvements}
\section{Conclusion}
\section{References}

\subsection{Purpose}
Identify the purpose of this SDD and its intended audience. (e.g. This software design
document describes the architecture and system design of XX.. ).

\subsection{Scope}
Provide a description and scope of the software and explain the goals, objectives and benefits
of your project. This will provide the basis for the brief description of your product.

\subsection{Overview}
Provide an overview of this document and its organization.

\subsection{Reference Material}
This section is optional.
List any documents, if any, which were used as sources of information for the test plan.

\subsection{Definitions and Acronyms}
This section is optional.
Provide  definitions of all terms, acronyms, and  abbreviations that might exist to  properly 
interpret the SDD. These definitions should be items used in the SDD that are most likely not 
known to the audience.

\begin{table}[h]
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Term}} & \multicolumn{1}{c|}{\textbf{Definition}}                                                                                              \\ \hline
Software Design Document (SDD)      & \begin{tabular}[c]{@{}l@{}}Used as the primary medium for communicating software \\ design information.\end{tabular}                  \\ \hline
Design Entity                       & \begin{tabular}[c]{@{}l@{}}An element of a design that is structurally and functionally \\ distinct from other elements.\end{tabular} \\ \hline
\end{tabular}
\end{table}

\section{System Overview}
Give a general description of the functionality, context and design of your project. Provide any 
background information if necessary.

\section{System Architecture}
\subsection{Architectural Design}

\begin{figure}[tbh]
\centering
\includegraphics[width=0.7\linewidth]{./image}
\caption{Architectural Design}
\label{fig:image}
\end{figure}

\subsection{Decomposition Description}

\subsection{Design Rationale}

\section{Data Design}
\subsection{Data Description}

\subsection {Data Dictionary}

\section{Component Design}

\section{Humnan Interface Design}

\subsection {Overview of User Interface}

\subsection {Screen Images}

\subsection {Screen Objects and Actions}

\section{APPENDICES}

\section {References}

\end{document}

### 2025-04-25 TechDoc/TechDoc.tex
Adding barcoding work to TechDoc

@@
 \subsubsection{User Interface}
+\paragraph{Feature Description}
+Given input parameters specified by the user, barcodes are generated via the backend and automatically downloaded for the user in txt format.
+The downloaded file is titled in the following format for easy identification and to prevent files being overwritten: \textit{SystemDateAndTime\_TB\_Barcode\_Gen\_AccessionName\.txt}.
+
+The file contains the following information: 
+\begin{itemize}
+    \item \textbf{Barcode sequence} - The base with the accession-specific SNP is enclosed in square brackets
+    \item \textbf{Barcode Start \& Barcode End} - The absolute position of the barcode in the chromosome 
+    \item \textbf{Variant Count} - The number of accession-specific SNPs that are present in the barcode 
+    \item \textbf{Variant Position} - The relative position of the accession-specific SNPs in the barcode 
+    \item \textbf{Repeat Sequence} - The sequence of the 2-bp repeating region. A repeating region is defined as 2 base pairs repeating 3+ times 
+    \item \textbf{Repeat Multiplier} - The number of times the repeat sequence is repeated 
+    \item \textbf{Repeat Start-End} - the relative start and end position of the repeat region within the barcode sequence 
+    \item \textbf{GC Content} - The GC content of the barcode 
+\end{itemize}
+
 \subsubsection{Technical Description}
+
+Barcode generation is controlled by two scripts: \verb+barcode_finder.sh+ and \verb+find_barcode.py+.
+
+\paragraph{Calling Tersect CLI to extract variants}: 
+\verb+Barcode_finder.sh+ is run on the command line, and takes as input accession name, chromosome, interval start position, interval end position, barcode size, maximum variant number, reference fasta, and tersect TSI index.
+The chromosome and interval start and end positions are used to define the searchable \verb+REGION+, and the accession name is used to define \verb+SAFE_ACC+ which is used to save files to a temporary file.
+
+The tersect CLI command \verb+tersect view+  is called to extract all variants within the specified region for the specified accession.
+The output is saved as a temporary TSV file as: \verb+${SAFE_ACC}_acc_unique.tsv+.
+The tersect CLI command tersect view is again used to extract all variants within the specified region for all accessions except the specified accession.
+This output is saved as a temporary TSV file as: \verb+${SAFE_ACC}_union_vars.tsv+.
+
+Lastly, the \verb+find_barcode.py+ file is called, passing as arguments the accession name, reference fasta, chromosome, interval start position, interval end position, barcode size, maximum variant number, and the tersect output files  \verb+${SAFE_ACC}_acc_unique.tsv+ and  \verb+${SAFE_ACC}_union_vars.tsv+.
+
+\paragraph{Generating barcodes}: 
+\verb+Find_barcode.py+ requires the following dependencies: argparse, SeqIO, and datetime.
+
+The reference fasta is parsed using \verb+SeqIO.parse+, yielding sequence records that are converted to a dictionary using \verb+SeqIO.dict+.
+Using the user-inputted chromosome name, \verb+[args.chrom].seq+ extracts the chromosome sequence from the dictionary and saves it to the variable ref.
+The user-defined interval start and end positions are used to extract the \verb+ref_window+ from ref.
+
+The \verb+load_variant_file()+ method imports the tersect output files and creates a dictionary, with variant position being stored as the key and a tuple containing the original base and the alternate base being stored as the dictionary value.
+Then, the \verb+remove_overlapping_variants()+ method compares the dictionaries and creates a new dictionary \verb+new_unique_vars+ with the same format, containing variants that are only present in the specified accession and not also present in any other accession.
+The method \verb+apply_variants_to_sequence()+ then uses the \verb+ref_window+ and \verb+new_unique_vars+ to generate an accession-specific sequence, \verb+unique_seq+, containing accession-specific SNPs.
+Lastly, the \verb+find_barcode_windows()+ method compares \verb+unique_seq+ against \verb+ref_window+ and using a sliding window of 1 base pair, identifies regions where the two sequences differ.
+The sequence, along with the absolute start and end position, is saved to the variable barcodes.
+
+\paragraph{Output file and barcode stats}: 
+Statistical metrics are calculated for each barcode, using custom methods.
+The number of accession-specific variants is calculated using \verb+count_variant_number()+, which also records variant position within the barcode.
+The variant position is used to highlight the variants within the barcode using the custom \verb+highlight_ref_alt_positions()+ method, which encloses the variant in the following format: \verb+[original base/alternate base]+.
+Repeat content is calculated using \verb+find_dinucleotide_repeats_custom()+, with repeats being defined as regions where a dinucleotide (2-base pair) sequence repeats consecutively three or more times.
+The repeating dinucleotide, number of times the dinucleotide repeats, and the start and end positions of the repeat region within the barcode are returned.
+GC content is calculated as a percentage to six decimal places using \verb+calculate_gc_content()+.
+
+The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+'TB_Barcode_Gen'+, and the specified accession.
+
 \subsubsection{Test Results}
 \subsubsection{Future Improvements}
-\subsection{Automated Introgression Search Extension}
-\subsubsection{Rationale}
-\subsubsection{User Interface}
-\subsubsection{Technical Description}
-\subsubsection{Test Results}
-\subsubsection{Future Improvements}
-\section{Conclusion}
+\subsection{Future Work and Extension Design}
+\subsubsection{Automated Introgression Search Extension}
 \section{References}

### 2025-04-26 Article/Article.tex
Structure Update

@@
 
-\title[Short Article Title]{TersectBrowser+ : A single browser for in-depth introgression analysis of resequenced genome datsaets}
+\title[Short Article Title]{TersectBrowser+ : A single browser for in-depth introgression analysis of re-sequenced genome datesets}
 
@@
 
+% Variant comparison important, tools are slow, what Tersect does.
+Tersect produced because tools slow.
+We modify Tersect creating Tersect+ to bringing additional functionality to aid the investigation of cultivar introgressions for biologists.
+
 \abstract{The field of cultivar genomics has evolved rapidly, with growing emphasis on mapping introgressions—segments of DNA from wild plant species that have been incorporated into cultivated lines.
@@
 
+% # Intro / Current Situation & Problem & Tersect Solution
+% Large-scale resequencing projects enable comparative genomic studies to identify key variants and structures like haplotypes or introgressions.
+studies rely on set theoretical operations, but current tools (e.g., BEDOPS, BCFtools, BEDTools) are limited in query complexity and speed.
+Tersect, using its own query language and fast execution through bitmap indices, enabling complex, real-time genome queries.
+This will become: a part on slowness 'As shown in [ref = Tom Fady paper] tools slow'.
+Then add something on functionality of these tools EITHER 'slow tools have functionality' or 'tools do not have functionality'.
+Then 'Tersect+ provides functionality.
+
 \section{Expanding landscape of plant genomic studies}\label{sec2}
@@
 Foundational work such as that by Aflitos et al.
-\cite{RN7} , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
+(2014) , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
 
@@
 
-\section{This is an example for first level head}\label{sec3}
+\section{Tersect Browser+}
 
-\subsection{This is an example for second level head - subsection head}\label{subsec2}
+% # Built based on Tersect Browswer... About Tersect About Tersect+- unified database, indexes and tracks and how speed is achieved. 
+% Builds a unified allele database from VCF files, using bitmap indexing to track presence/absence of variants.
+The system links variant positions to bitmap indices, allowing efficient region-specific searches and genome comparisons via compressed, sparse data structures.
+Indexing is fast and optimized for read-heavy workflows.
+While updating is less efficient, batch additions mitigate this.
+A custom parser converts set-based queries into abstract syntax trees (ASTs), which are then optimized and reused across queries for speed.
+We summarize here the functionality.
+Probably in the following structure.
+Browser; Models; Search; and Barcodes and Introgressions [if we get there.] We do this with a paragraph on each.
+We write similarly to the Fady and Tom email.
+We could do it in two parts:  Browser and Gene Model and Search and Barcodes including introgression search, we do that.
+The structure in paragraphs will be:
 
-\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}
+\section{Features}
 
-\paragraph{This is an example for fourth level head - paragraph head}
+% Browser  - all these will be kept short paragraphs.
+% Gene Models
+% Search 
+% Barcodes
+% [Introgression]
 
-\section{Conclusion}
+\section{Funding}
 
-Some Conclusions here.
+%\section{This is an example for first level head}\label{sec3}
+
+% \subsection{This is an example for second level head - subsection head}\label{subsec2}
+
+% \subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}
+
+% \paragraph{This is an example for fourth level head - paragraph head}
+
+% \section{Conclusion}
+
+% Some Conclusions here.
 
 %%%%%%%%%%%%%%
-
-\begin{appendices}
-
-\section{Section title of first appendix}\label{sec11}
-
-\subsection{Subsection title of first appendix}\label{subsec4}
-
-\section{Example of another appendix section}\label{sec13}%
-
-\end{appendices}
 
@@
 
-\section{Competing interests}
+% \section{Competing interests}
 
-\section{Author contributions statement}
+% \section{Author contributions statement}
 
-\section{Acknowledgments}
+% \section{Acknowledgments}
 
-\bibliography{examplerefs}
-
-\begin{thebibliography}{10}
-\bibitem{bahdanau2014neural}
-Dzmitry Bahdanau, Kyunghyun Cho, and Yoshua Bengio.
-\newblock Neural machine translation by jointly learning to align and
-  translate.
-\newblock {\em arXiv preprint arXiv:1409.0473}, 2014.
-\end{thebibliography}
+\bibliographystyle{plain}
+\bibliography{Article/article_ref}
 

### 2025-04-26 Article/Article.tex

@@
 Foundational work such as that by Aflitos et al.
-(2014) , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
+\cite{RN7} , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
 

### 2025-04-26 Article/Article.tex
Draft introduction added with references

--- 2025-04-26 Article/Article.tex
@@
 
+% With Tom first, as first author and Students in alphabetical order. 
+\author[1,$\ast$]{Tomasz Kurowski\ORCID{}}
+\author[2]{Gabrielle Baumberg}
+\author[2]{Gregory Lupton}
 \author[2]{Tanya Stead}
 \author[2]{David Oluwasusi}
-\author[2]{Gregory Lupton}
-\author[2]{Gabrielle Baumberg}
-\author[1,$\ast$]{Tomasz Kurowski\ORCID{}}
 
@@
 
-\section{Expanding landscape of plant genomic studies}\label{sec2}
+Cultivar genomics has evolved rapidly, with growing emphasis on mapping introgressions, segments of DNA from wild plant species that have been incorporated into cultivated lines.
+A variety of bioinformatic tools aim to interpret this expanding complexity of plant genomic data, predict introgressions, and identify commercially relevant regions for future breeding (Qin et al., 2021)\cite{Qinetal2023}.
+Demand for tailored tools to explore these complex genomic queries in real-time is increasing.
 
-The field of tomato genomics has continued to evolve rapidly, with a growing emphasis on mapping introgressions — segments of DNA from wild tomato species that have been incorporated into cultivated lines.
-Foundational work such as that by Aflitos et al.
-(2014) , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
+As shown previously by Kurowski & Mohareb (2020)\cite{KurowskiMohareb2019}, Tersect is a lightweight, high-performing, real-time genome querying command-line utility that uses its own query language and bitmap indices to interpret and apply flexible set theoretical expressions to sets of sequence variant data for tomato genomes.
+In comparison to other tools such as BEDOPS, BCFtools and BDETools (Quinlan & Hall, 2010\cite{Qinlan2010}; Danecek et al., 2011)\cite{}, Tersect performs from three to over a hundred times faster than BCFTools (Kurowski & Mohareb, 2020)\cite{KurowskiMohareb2019}.
+
+For biologists, command-line-based tools can be cumbersome.
+Web-app based tools with intuitive data visualization, real-time querying, and introgression mapping provide much better user experiences and better meet the requirements for interactive exploration of large-scale variant datasets and immediate visual feedback on genomic regions of interest.
+
+In 2022, a web-based version of Tersect, Tersect Browser, was created (Kurowski, 2023\cite{KurowskiphdThesis}) using \boldsymbol{Include dataset}.
+Tersect Browser offered multiple front-end extension capabilities that allowed user to view and analyse data without leaving the page.
+These include: a dynamic introgression heatmap, tree navigation of accessions, and region-focused querying of genomic variants.
+
+The field of tomato genomics, however, is growing with increasing emphasis on mapping introgressions and higher volumes of genomic data.
+As the field has built on foundational work of Aflitos et al, (2014)\cite{Aflitos2014}, which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis, there is increasing requirements for more powerful and more capable introgression visualization and comparative analysis tools.
 
 Recent studies have expanded on this base with increasingly sophisticated tools and datasets.
+Notably, Liu et al.
+(2023)\cite{liu_effects_2023} used tomato pan-genomes built from hundreds of accessions to uncover regions of structural variation and introgression hotspots, many of which affect key traits like fruit size, shape, disease resistance, and flavour.
+The construction of these pan-genomes enables researchers to observe variation absent in the reference genome, improving trait mapping and introgression detection.
+
+Here, we present Tersect Browser+, an updated introgression analysis platform with expanded front-end features extension capabilities that users can view and analyze data without leaving the page.
+It integrates JBrowse (Skinner et al., 2009\cite{skinner_jbrowse_2009}; Buels et al., 2016 \cite{buels_jbrowse_2016}) capabilities to capture gene models, exon-intron structures, and positional gene information for biologists inspecting tomato introgression data sets.
+
+JBrowse is a widely used, web-based genome browser designed for fast, interactive visualization of genomic data.
+Developed as a successor to earlier tools like GBrowse, JBrowse enables users to view and explore large-scale genome annotations, sequence data, and variation information through a modern, dynamic interface (Skinner et al, 2009 \cite{skinner_jbrowse_2009}).
+It supports a wide variety of standard file formats, including GFF, BAM, VCF, and FASTA, and is highly extensible through plugins, making it adaptable for a range of research applications from comparative genomics to functional annotation, including introgression web-app based analysis (Buels et al., 2016\cite{buels_jbrowse_2016}).
+Researchers often deploy JBrowse to visualize genome assemblies, transcriptomes, and variant information.
+Its active development and ease of integration into web servers have made it a common choice for bioinformatics platforms.
+
+Tersect Browser+ also includes other features such as a search feature and barcode generating feature.
+It allows for the investigation of other crop datasets, for example Soybean [include specific species type/name].
+We identify previously published introgressions with our tool, illustrating its potential for future research, and enriching plant breeders' understanding of crop cultivar genomics.
+
+%\section{Expanding landscape of plant genomic studies}\label{sec2}
+
+% As previously shown by Kurowski & Mohareb (2020), Tersect, a lightweight, high-performing, real-time genome querying command-line utility that uses its own query language and fast execution through bitmap indices to interpret and apply flexible set theortical expressions to sets of sequence variant data, allows [something] identification of key variants and structures like haplotypes and introgressions in support of comparative genomic studies.
+Tersect returned
+
+% The field of tomato genomics has continued to evolve rapidly, with a growing emphasis on mapping introgressions — segments of DNA from wild tomato species that have been incorporated into cultivated lines.
+Foundational work such as that by Aflitos et al.
+\cite{RN7} , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
+
+% Recent studies have expanded on this base with increasingly sophisticated tools and datasets.
 Notably, Liu et al.
@@
 
-\subsection{Background on introgression studies}\label{subsec1}
+% \subsection{Background on introgression studies}\label{subsec1}
 
-The role of introgression in water-tolerance Kubond et al. (2022), shows how such analysis will be important in identifying variant resistance to climate change.
+% The role of introgression in water-tolerance Kubond et al. (2022), shows how such analysis will be important in identifying variant resistance to climate change.
 
-To manage and interpret the expanding complexity of genomic data, a wide range of bioinformatic tools have been deployed.
+% To manage and interpret the expanding complexity of genomic data, a wide range of bioinformatic tools have been deployed.
 Established platforms such as SnpEff, VEP, BEDTools, BEDOPS, CrossMap, and the UCSC Genome Browser continue to support annotation, variant effect prediction, and coordinate translation.
@@
 
-Meanwhile, long-read sequencing technologies (e.g., PacBio, Hi-C, and optical mapping) as applied in SL4.0 (Hosmani et al., 2019), and innovations in in situ hybridization (Shearer et al., 2014), have markedly improved the resolution of reference genomes—critical for detecting subtle and ancient introgressions.
+% Meanwhile, long-read sequencing technologies (e.g., PacBio, Hi-C, and optical mapping) as applied in SL4.0 (Hosmani et al., 2019), and innovations in in situ hybridization (Shearer et al., 2014), have markedly improved the resolution of reference genomes—critical for detecting subtle and ancient introgressions.
 
-Beyond tomato, research in soybean introgression offers valuable cross-crop insights.
+% Beyond tomato, research in soybean introgression offers valuable cross-crop insights.
 Studies have shown how introgressed genomic regions from wild Glycine soja into Glycine max improve traits like drought resistance, seed oil composition, and disease resilience.
@@
 
-Together, these advances paint a dynamic picture of tomato genomics: one where introgression is not only a window into domestication and evolutionary history but also a powerful mechanism for ongoing improvement, including against threats such as climate change.
+% Together, these advances paint a dynamic picture of tomato genomics: one where introgression is not only a window into domestication and evolutionary history but also a powerful mechanism for ongoing improvement, including against threats such as climate change.
 The fusion of high-resolution genomic data, computational methods, and specialized tools enables researchers to explore tomato’s genetic landscape with increasing clarity and sophistication.
-(Introgression hybridisation is an important factor in crop improvement.
+% (Introgression hybridisation is an important factor in crop improvement.
 Biologists need to determine introgressions and identify donor species through visualising genetic distance and phylogenetic relationships based on the whole genome variant data.
@@
 
-\section{Tersect Browser+}
+\section{Features}
+
+Tersect Browser+ is a web application based on Tersect Browser
+
+\subsubsection{Gene Browser}
+
+We will include text here on the feature. I will look through the Technical Documentation to put this together.
+
+\subsubsection{Gene Model}
+\subsubsection{Search Features}
+\subsubsection{Barcode Generator}
 
@@
 
-\section{Features}
-
 % Browser  - all these will be kept short paragraphs.
@@
 \section{Funding}
+
+A question for Tom - we see this included in many papers published in this journal. What should we include?
 
@@
 \section{Code availability}
-The technical documentation is available at XX. The source code for the project is freely accessible on GitHub at XX.
+
+We will add in a link to the github reference in here and a link to the technical documentation.
 
@@
 \bibliographystyle{plain}
-\bibliography{Article/article_ref}
+\bibliography{Article/article_references}
 

### 2025-04-28 Article/Article.tex
Including Feature section intro.

@@
 \section{Features}
+Tersect Browser+ is a web application that depends on resequenced data in VCF format and uses a reference genome in FASTA format.
+Utilizing a MEAN stack (MongoDB, Express, Angular, and Node) architecture, it creates a webpage displaying heatmaps and phylogenetic trees for introgressions within the loaded crop dataset.
+This provides an interface for users to identify differences across accessions within the same species dataset.
+It also includes a JBrowse interface for investigating specific variants, displaying selected genes directly alongside the generated heatmap.
+A search feature highlights high-impact variants on the webpage, helping to distinguish the selected reference genomes.
+Tersect Browser+ also offers barcode generation functionality for various cultivars and specific species in the dataset.
 
-Tersect Browser+ is a web application based on Tersect Browser
+\begin{figure} [http]
+  \centering
+  \includegraphics[width=\textwidth]{path/to_image.png}
+  \caption{Tersect Browser+ overview. The browser provides a single way to analyse variant data without leaving the page.[More description]}
+  \label{fig:your_label}
+\end{figure}
 
-\subsubsection{Gene Browser}
+\subsubsection{Gene Brwoser}
 
@@
 \subsubsection{Gene Model}
+
+We will include text here on the feature. I will look through the Technical Documentation to put this together.
+
 \subsubsection{Search Features}
+
+We will include text here on the feature. I will look through the Technical Documentation to put this together.
+
 \subsubsection{Barcode Generator}
+
+We will include text here on the feature. I will look through the Technical Documentation to put this together.
 

### 2025-04-28 TechDoc/TechDoc.tex
Update TechDoc.tex

@@
 \title{
-{\includegraphics[width=3cm, height=2.5cm]{./Cran.jpg}}
+{\includegraphics[width=3cm, height=2.5cm]{TechDoc/Cran.jpg}}
 \\
-\includegraphics[width=7cm, height=1cm]{./TB.jpg}
+\includegraphics[width=7cm, height=1cm]{TechDoc/TB.jpg}
 \\
@@
 
-\author{David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}
+\author{Extra author, David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}
 

### 2025-04-28 TechDoc/TechDoc.tex
Update TechDoc.tex
Added barcoding section

@@
 \title{
-{\includegraphics[width=3cm, height=2.5cm]{TechDoc/Cran.jpg}}
-\\
-\includegraphics[width=7cm, height=1cm]{TechDoc/TB.jpg}
-\\
+%{\includegraphics[width=3cm, height=2.5cm]{TechDoc/Cran.jpg}}
+%\\
+%\includegraphics[width=7cm, height=1cm]{TechDoc/TB.jpg}
+%\\
 {Technical Documentation for TersectBrowser+}
@@
 This Technical Documentation describes the particular details of updating the original TersectBrowser developed by Tomasz Kurowski into the TersectBrowser+ by integrating multiple extension features.
+
 \section{Architecture}
@@
 
+For the TersectBrowser+ update to the original TersectBrowser, main changes have been made to the frontend implementation of features added in a separate extension section of the project.
+The main addition is in a 'genome-browser' extension, which contains elements for each of the specific features mentioned in the project brief.
+
 \section{Deployment}
+The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment, however the main outline is given here.
 
@@
 
-\section{TersectBrowser}
+The main requirements for TersectBrowser+ deployment are:
+\begin{itemize}
+    \item nvm (versions 16 and 22 specifically, as these will be deployed separately)
+    \item npm registery \url{https://registry.npmjs.org/}
+    \item Angular CLI v1.7.1
+    \item Tersect CLI
+    \item JBrowse CLI
+    \item MongoDB
+    \item Python virtual environment
+    \item RapidNJ / Rosetta
+\end{itemize}
+
 \section{TersectBrowser+}
@@
 
-\begin{figure}[tbh]
-\centering
-\includegraphics[width=0.7\linewidth]{./image}
-\caption{Architectural Design}
-\label{fig:image}
-\end{figure}
-
 \subsection{Decomposition Description}

### 2025-04-28 Article/Article.tex
Adding rationale to extension intros

@@
 Cultivar genomics has evolved rapidly, with growing emphasis on mapping introgressions, segments of DNA from wild plant species that have been incorporated into cultivated lines.
-A variety of bioinformatic tools aim to interpret this expanding complexity of plant genomic data, predict introgressions, and identify commercially relevant regions for future breeding (Qin et al., 2021)\cite{Qinetal2023}.
+A variety of bioinformatic tools aim to interpret this expanding complexity of plant genomic data, predict introgressions, and identify commercially relevant regions for future breeding (Qin et al., 2021) \cite{Qinetal2023}.
 Demand for tailored tools to explore these complex genomic queries in real-time is increasing.

### 2025-04-28 TechDoc/TechDoc.tex
Adding rationale to extension intros

@@
 \usepackage{soul}
+\usepackage{listings}
 \usepackage[a4paper,width=150mm,top=25mm,bottom=25mm]{geometry}
@@
 
-\author{Extra author, David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}
+\author{David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}
 
@@
 
+\subsection{Dataset Upload}
+
+The Admin is defined as an individual with bioinformatics and software design knowledge, background comprehension for all sections of this Technical Documentation, and acting to facilitate the use of a deployed TersectBrowser+ by the plant breeder main user of the software.
+When a new dataset is requested for deployment on TersectBrowser+, the Admin will need to follow the below steps:
+\begin{enumerate}
+    \item Collect background context and required data from the user.
+\begin{itemize}
+    \item Reference genome in fasta format (may be compressed).
+Size and chromosome number will impact the speed of dataset upload and deployment - TersectBrowser+ has been tested with a genome size of ~1GB, and 20 chromosome pairs.
+    \item Resequenced genome dataset in VCF format (may be compressed, as a multi-sample VCF or as a directory of individual files).
+TersectBrowser+ has been tested with a dataset size of ~500 VCF files representing individual accessions.
+    \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
+    \item A GFF file of gene model information for the reference genome, produced using SNPeff software.
+If not using SNPeff, ensure that variant impact levels are categorised into 'High', 'Medium', and 'Low' impact.
+If not provided, the gene model track in the Variant Browser will be unavailable.
+    \item Whether there are prior identified introgressions known to be functionally relevant for the species.
+These can be used to 'check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
+\end{itemize}
+    \item Clone the GitHub repository of TersectBrowser+ to the local machine.
+Follow the README from \hyperlink {https://github.com/Tersect-Browser/Tersect-browser.git}{TersectBrowser+ GitHub} for installation instructions.
+    \item Download relevant files from section (1) to a new folder within the root directory of the local Tersect-browser repository.
+    \item Set environment variables \verb+fasta+, \verb+gff+, and \verb+vcfs+ according to their new location relative to the root of Tersect-browser.
+    \item Run the script \verb+setup_new_tbrowser_dataset.py+ using the environment variables of files as input:
+
+        \verb+python setup_new_tbrowser_dataset.py -f ${fasta} -g ${gff} -V ${vcfs}+
+        \begin{itemize}
+            \item After running the script, there should be visible a new \verb+config.json+ file in the root folder, as well as a new folder \verb+./~/mongo-data/gp_data_copy+ containing fasta files, gff files, and VCF files along with their index files.
+The main Tersect index file should be in the folder above.
+This is where the server will look to find information on tracks when generating the Variant Browser panel.
+        \end{itemize}
+\end{enumerate}
+
+specified location within the repository: \textit{Tersect-browser/\(\sim \)/mongo-data/gp\_data\_copy/}
+\begin{itemize}
+    \item Download all VCF files in the resequenced dataset. Run tabix on these files to create .tbi index files.
+    \item Download the reference genome FASTA file. Run SAMtools faidx to create .fai index file.
+    \item Download (or generate) an annotation GFF file for the resequenced dataset using SNPeff.  Index the GFF file. Run tabix on this file to create a .tbi index file.
+\end{itemize}
+\begin{itemize}
+\item Use JBrowse CLI text-index tool to generate Trix text searching files (.ix, .ixx, and .meta.json) from the GFF annotation file.
+\item Run the script \verb+add_config_tracks.py+ using these downloaded files. This will use a set of jbrowse arguments to load the datafiles as tracks in the Variant Browser component.
+\item Run the script \verb+change_dataset_files.py+ using these downloaded files as CLI arguments, in order to update the references to these data files in the project.
+\item Run the script \verb+add_example_dataset.sh+ to prepare the dataset.
+\end{itemize}
+
 \section{TersectBrowser+}
@@
 \subsubsection{Rationale}
+The main benefit of TersectBrowser relative to similar genome comparison software for displaying key differences within resequenced datasets is its use of Tersect CLI, to index and then quickly query these genomes for investigation at user-defined scale.
+However, in order for TersectBrowser+ to be competitive as a tool with other browsers, it was necessary to implement features that enhance further investigation within the scope of the same browser.
+The main view of the original TersectBrowser displayed the density of variants along each chromosome of the reference genome, a view compatible with the addition of a variant browser with multiple tracks providing extra information at a per-base level.
+
+After a review of node-compatible browser tools such as CRAMER, JBrowse was ultimately chosen to carry out this extension request.
+The initial build of JBrowse into TersectBrowser+ as an extension focused on retaining features most relevant to the user aims, and bringing the styling of the panel in line with the rest of the page.
+This extension also involved generating a popup window of the same JBrowse component to be initialised when the user was interested in a specific accession or bin region.
 
@@
         \item Variant tracks for each loaded accession
-        \item Gene models FeatureTrack (Ext-GM) 
+        \item Gene models FeatureTrack (\nameref{sec:Models})
     \end{itemize}
@@
 
+When a new dataset is added, the custom addition script automatically updates the names of tracks and assembly in the relevant files within the genome-browser extension.
+This flexibility allows the Variant Browser extension to be accessible by plant breeders working on many different datasets.
+TersectBrowser was initially validated with a human genome dataset, so it was important to maintain this flexibility of input in the new extensions.
+
 \subsubsection{Future Improvements}
-(Current) Limitations:
-
-Only SL2.50 loaded as assembly and reference
-
-Can only view variant tracks for 150\_VCF\_Tomato dataset - this is hard coded, and would need to be changed if we want this to be dynamic (when the user adds their own dataset) / if we as admins want to load soybean dataset!
 
@@
 \subsubsection{Rationale}
+Once the Variant Browser panel was added to the main page of TersectBrowser+, this facilitated the inclusion of a specified gene model track.
+This allows the plant breeder user to identify an area of high variant density on the heatmap, and then check the gene model track to investigate whether this region occurs near a genetic feature with functional relevance to the species.
+
 \subsubsection{User Interface}
@@
     \item In the Jbrowse window, the first track shows gene models based on the GFF file provided.  
-    \item Gene model annotations are based on the official \hl{tomato genome annotation v2.4}: \href{https://phytozome-next.jgi.doe.gov/info/Slycopersicum_ITAG2_4#:~:text=Genome%20Overview,genome%2Cwith%2034%2C725%20gene%20models}{Slycopersicum ITAG annotation v2.4}
+%    \item Gene model annotations are based on the official \hl{tomato genome annotation v2.4}: \href{https://phytozome-next.jgi.doe.gov/info/Slycopersicum_ITAG2_4#:~:text=Genome%20Overview,genome%2Cwith%2034%2C725%20gene%20models}{Slycopersicum ITAG annotation v2.4}
+% The above is tomato-specific. Not sure where it fits
     \item The popup view of the same Jbrowse component has similar default: the first track is the gene models track, and the second track is the variant track for the selected accession.
@@
 \subsubsection{Test Results}
+This extension has been tested with the additional Soybean dataset, with correct functionality.
+The relevance of the gene models in the GFF file depend on what the user provides in the dataset, and how the GFF file was created.
+
 \subsubsection{Future Improvements}
@@
 \subsubsection{Rationale}
-The aim of this extension is to allow the user to search specific genes and identify which accession contains a high/medium/low impact variant impacting that gene.
+The aim of this extension is to allow the user to search specific genes and identify which accession contains a \hl{high/medium/low} impact variant affecting that gene.
+The bins containing high impact variants for that gene will be highlighted at the accession level, allowing high resolution investigation of variants.
+The user may alternatively search for any high impact variants within a specified region, in case they lack a specific gene name to search.
+The highlighted bins may then be selected and opened in the popup \nameref{sec:Browser} window for further investigation.
+
 \subsubsection{User Interface}
@@
 \subsubsection{Future Improvements}
+
 \subsection{Barcode Generation Extension}
 \subsubsection{Rationale}
+A key aim for plant breeders when comparing their own plant strains with other cultivars in resequenced datasets is to be able to identify their strain at the genome level.
+This would allow for protection of intellectual property in the case of an advantageous new introgression that can be uniquely identified.
+The Barcoding extension allows this unique identification by generating a range of barcodes with a user-specified maximum number of SNPs, so that primers for this barcode are easily generated in a wetlab environment.
+Metrics comparing these putative barcodes are also provided to the user, and available for local download.
+
 \subsubsection{User Interface}
@@
 
-\section{System Overview}
-Give a general description of the functionality, context and design of your project. Provide any 
-background information if necessary.
-
 \section{System Architecture}
@@
 
-\subsection{Decomposition Description}
-
 \subsection{Design Rationale}
 
-\section{Data Design}
 \subsection{Data Description}
@@
 \subsection {Data Dictionary}
-
-\section{Component Design}
-
-\section{Humnan Interface Design}
-
-\subsection {Overview of User Interface}
-
-\subsection {Screen Images}
-
-\subsection {Screen Objects and Actions}
 

### 2025-04-28 UserManual/UserMan.tex
Adding rationale to extension intros

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{soul}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm]{geometry}

\title{
{\includegraphics[width=3cm, height=2.5cm]{Cran.jpg}}
\\
\includegraphics[width=7cm, height=1cm]{TB.jpg}
\\
{User Manual for TersectBrowser+}
}

\author{David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}

\begin{document}
\sloppy % allows line spacing stretch to justify the text

\maketitle

\section{Overview}
TersectBrowser+ is designed to be useful to a range of Users. Hosting features within the same browser allows the conducting of full analyses within the same page, benefitting both a more experienced researcher as well as a novice to the field.

\section{Use Cases - Video transcript}

\subsection{Example 1:} A tomato genetics researcher wishes to identify differences between a domestic tomato genome and a large number of resequenced genomes of wild relatives.

\paragraph{User profile:} Extensive knowledge of introgression genetics and in particular the tomato genome and different species. Lacks knowledge of software development and the different file types involved in setting up the software.

\begin{enumerate}
  \item The researcher reaches out to administration or informatics support to set up an instance of TersectBrowser+ that contains the large resequenced dataset and relevant GTF files, Reference genomes, and Metadata (e.g. geographic location of species) annotations for the difference accessions. 
  \item The researcher uses a Local Host Web Address to access TersectBrowser+. The default view is a list of all accessions in the dataset, with one chromosome view selected and a set bin size and zoom level. 
  \subitem  At this stage, the researcher can see a heatmap of SNP density across the different accessions, and gain a brief overview of which accessions differ most from the selected reference sequence. The researcher also can see which locations on the chromosome contain the most variation in SNP density. 
  \item The researcher is particularly interested in a dark section of the heatmap (I.e. high density of SNPs relative to the selected reference accession), that lies across a chromosomal region containing genes involved in heat stress regulation. The researcher uses the click and select feature along the scale bar to bring up a popup message with further options, and selects the "Set as interval" feature. 
  \subitem At this stage the researcher understands that the view will change so that the selected chromosomal region (containing high density SNPs in certain accessions) will be zoomed to in some way.  
  \subitem The researcher sees that TersectBrowser+ view zooms to the selected region. The variation in grey bin highlighting is more obvious in this view. The accession names are also zoomed in, and therefore easier to read.
  \item The researcher now sees that a particular set of 5 accessions are the darkest on the heatmap. The researcher would like to know if any key genes cover the dark regions for these 5 accessions. They understand that the gene browser containing gene models and the reference genome is accessible by clicking the binoculars button.  
  \subitem The researcher sees a new view with the heatmap shifted down so there is a panel of the gene browser at the top of the page. The default view of the gene browser panel is the gene models track, and the first two accession vcf files in the resequenced dataset. The researcher clicks the track selecter button to search for the names of the 5 particular accessions. 
  \subitem The researcher finds that the zoom on the heatmap has a minimum level past which the scroll zoom no longer works. However, the gene browser panel can be zoomed further than this by selecting a region on the scale bar and clicking "Zoom to region". 
  \subitem This then zooms to a view where the gene models are visible alongside the variants specific to each accession.  
  \item The researcher finds a gene model that they know to be important for heat stress. Upon clicking this gene in the browser \hl{panel,}
\end{enumerate}

\subsection{Example 2:} A plant breeder intends to gain intellectual property (IP) recognition for a particular trait of their Soybean plant breed, and wants to find quantitative evidence to support their application using the genome of this breed.

\paragraph{User profile:} High familiarity with phenotypic traits of different Soybean plant breeds, as well as history of intended Soybean introgressions by different plant breeders. Lacks knowledge of statistical generation of phenetic trees using pairwise relationships, as well as how the data in the TersectBrowser+ is set up to view the whole dataset.

The breeder opens up the phenetic tree view, in order to view a summary of pairwise differences between the different accessions.

The breeder uses the barcode search function to identify whether the sequence within the identified introgression in their plant breed can be used to uniquely identify that breed relative to the rest of the resequenced dataset.

The breeder exports the results of the barcode search to use in their application for IP recognition of their breed.

\subsection{Example 3:} A bioinformatics expert intends to learn more about what introgression means in practice as relating to differences between genomes of similar species.

\paragraph{User profile:} Expert knowledge in GUI Web-app design, as well as

The bioinformatician can zoom in the view using the plus/minus buttons on the home bar, or using the mouse zoom scroll (up and down).

The bioinformatician uses the introgression prediction tool to identify which high SNP density regions meet the threshold requirements for high likelihood of introgression from a wild cultivar into the domestic cultivar.

\section{Admin setup of browser}
The Admin is defined as an individual with bioinformatics and software design knowledge, background comprehension for all sections of this Technical Documentation, and acting to facilitate the use of a deployed TersectBrowser+ by the plant breeder main user of the software. When a new dataset is requested for deployment on TersectBrowser+, the admin will need to follow the below steps.

\subsection{Download of required material}
\begin{enumerate}
  \item Collect background context and required data from the user.
\begin{itemize}
  \item Reference genome in fasta format (may be compressed). Size and chromosome number will impact the speed of dataset upload and deployment - TersectBrowser+ has been tested with a genome size of ~1GB, and 20 chromosome pairs.
  \item Resequenced genome dataset in VCF format (may be compressed, as a multi-sample VCF or as a directory of individual files). TersectBrowser+ has been tested with a dataset size of ~500 VCF files representing individual accessions. 
  \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
  \item A GFF file of gene model information for the reference genome, produced using SNPeff software. If not using SNPeff, ensure that variant impact levels are categorised into 'High', 'Medium', and 'Low' impact. If not provided, the gene model track in the Variant Browser will be unavailable. 
  \item Whether there are prior identified introgressions known to be functionally relevant for the species. These can be used to 'check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
\end{itemize}
  \item Clone the GitHub repository of TersectBrowser+ to the local machine. Follow the README from \hyperlink {https://github.com/Tersect-Browser/Tersect-browser.git}{TersectBrowser+ GitHub} for installation instructions.
  \item Download relevant files from section (1) to a new folder within the root directory of the local Tersect-browser repository.
  \item Set environment variables \verb+fasta+, \verb+gff+, and \verb+vcfs+ according to their new location relative to the root of Tersect-browser.
  \item Run the script \verb+setup_new_tbrowser_dataset.py+ using the environment variables of files as input:

      \verb+python setup_new_tbrowser_dataset.py -f ${fasta} -g ${gff} -V ${vcfs}+
      \begin{itemize}
          \item After running the script, there should be visible a new \verb+config.json+ file in the root folder, as well as a new folder \verb+./~/mongo-data/gp_data_copy+ containing fasta files, gff files, and VCF files along with their index files. The main Tersect index file should be in the folder above. This is where the server will look to find information on tracks when generating the Variant Browser panel.
      \end{itemize}
\end{enumerate}

\section{Config file generation}
There are multiple extensions incorporated within TersectBrowser+, and all config files must be correctly generated for full functionality of the browser. However, each extension has different config requirements and may allow limited functionality before the generation of all config files.

\subsection{Variant Browser Pane}
A config.json file containing a/the genome assembly is first generated. 
\paragraph{Using JBrowse CLI}\mbox{}
\\\\
The fasta file must first be indexed as follows (this step can be omitted if a .fai file already exists):
\lstset{
  language=bash,
  basicstyle=\ttfamily
}
\begin{lstlisting}
    $ cd Tersect-browser/~mongo-data/gp_data_copy/
    $ samtools faidx genome.fa 
\end{lstlisting}
The config.json file is created and the genome assembly added to it via the following command:
\begin{lstlisting}
    $ jbrowse add-assembly <genome.fa> --load copy --out config.json
\end{lstlisting}
The variant tracks can then be added to the \verb+config.json+ file via the following commands. Jbrowse CLI expects the VCFs to be compressed with bgzip and indexed using tabix. 
\begin{lstlisting}
    $ bgzip file.vcf 
    $ tabix file.vcf.gz 
    $ jbrowse add-track file.vcf.gz --load copy --out config.json 
\end{lstlisting}
The generated config file will contain both the \verb+assembly[]+ and the \verb+tracks[]+. These can then be manually copied and pasted into their respective assembly.ts and tracks.ts files - the reference genome fasta will be added to the \textit{extension/genome-browser/src/app/react-components/assembly.ts} file, while the GFF file and all accession VCF tracks will be added to the \textit{extension/genome-browser/src/app/react-components/tracks.ts} file.

Correct config specifications for the reference genome assembly can be found \textunderscore{\hyperlink{https://jbrowse.org/jb2/docs/config_guides/assemblies/}{here}}.

\paragraph{Manual Generation}\mbox{}
\\\\
If Jbrowse CLI is not compatible with the machine, the assembly config can be generated by adding the following text, and \hl{changing the parameters trackID, name, and URI where necessary}.

\begin{lstlisting}
    $ cd extension/genome-browser/src/app/react-components/
    $ nano config.json

    "name": "SL2.50", 
    "sequence": { 
      "type": "ReferenceSequenceTrack", 
      "trackId": "SL2.50-ReferenceSequenceTrack", 
      "adapter": { 
        "type": "IndexedFastaAdapter", 
        "fastaLocation": { 
          "uri": "http://localhost:4200/TersectBrowserGP/tbapi/datafiles//SL2.50.fa", 
          "locationType": "UriLocation" 
        }, 
        "faiLocation": { 
          "uri": "http://localhost:4200/TersectBrowserGP/tbapi/datafiles//SL2.50.fa.fai", 
          "locationType": "UriLocation" 
        } } } 
\end{lstlisting}

The track config can also be generated by adding the following text, and \hl{changing the parameters trackID, name, and uri where necessary}.

\begin{lstlisting}
    { 
    "type": "VariantTrack", 
    "trackId": "S.lyc LA2706", 
    "name": "S.lyc LA2706", 
    "adapter": { 
        "type": "VcfTabixAdapter", 
        "vcfGzLocation": { 
            "uri": "http://localhost:4200/TersectBrowserGP/tbapi/datafiles/RF_001_SZAXPI008746-45.vcf.gz.snpeff.vcf.gz", 
            "locationType": "UriLocation" 
        }, 
        "index": { 
            "location": { 
                "uri": "http://localhost:4200/TersectBrowserGP/tbapi/datafiles/RF_001_SZAXPI008746-45.vcf.gz.snpeff.vcf.gz.tbi", 
                "locationType": "UriLocation" 
            }, 
            "indexType": "TBI" 
        } 
    }, 
    "assemblyNames": [ 
        "SL2.50" 
    ]      } 
\end{lstlisting}

A script to automate this process (the tracks will be more numerous than assemblies) can be found on \hl{Elvis} at \textit{gp\_data\_copy/scripts/generate\_track\_config.py}, and  a script to generate tbi files for all VCF files, \textit{generate\_tbi.sh}, is also present in this folder.

\subsection{Gene Model Track}
\paragraph{Using JBrowse CLI}\mbox{}
\\\\
Once a \verb+config.json+ file for the main reference assembly has been created, run the following commands to sort the GFF3 file, create the .tbi file, and add the config files to the \verb+config.json+ using these Jbrowse commands:
\lstset{
  language=bash,
  basicstyle=\ttfamily
}
\begin{lstlisting}
    $ jbrowse sort-gff <yourfile.gff> | 
        bgzip > <yourfile.sorted.gff.gz>
    $ tabix <yourfile.sorted.gff.gz>
    $ jbrowse add-track <yourfile.sorted.gff.gz> --load copy
\end{lstlisting}
This will create a \verb+FeatureTrack+ under the \verb+tracks[]+ section in the \verb+config.json+. Then copy and paste the \verb+FeatureTrack+ to the \textit{extension/genome-browser/src/app/react-components/tracks.ts} file.

Note: the Jbrowse "sort-gff" command just automates the following shell command 
\begin{lstlisting}
    $ (grep "^#" in.gff; grep -v "^#" in.gff | 
        sort -t"`printf '\t'`" -k1,1 -k4,4n)  
        > sorted.gff; 
\end{lstlisting}

\paragraph{Manual Generation}\mbox{}
\\\\
If Jbrowse CLI cannot be accessed, the \verb+FeatureTrack+ can be generated by adding the following text, replacing trackID, name, and uri with the respective inputs:
\begin{lstlisting}
    { 
        "type": "FeatureTrack", 
        "trackId": "ITAG2.4 Gene Models", 
        "name": "ITAG2.4 Gene Models", 
        "adapter": { 
          "type": "Gff3TabixAdapter", 
          "gffGzLocation": { 
            "uri": "http://localhost:4200/TersectBrowserGP/tbapi/datafiles/ITAG2.4_gene_models.sorted.gff3.gz", 
            "locationType": "UriLocation" 
          }, 
          "index": { 
            "location": { 
              "uri": "http://localhost:4200/TersectBrowserGP/tbapi/datafiles/ITAG2.4_gene_models.sorted.gff3.gz.tbi", 
              "locationType": "UriLocation" 
            }, 
            "indexType": "TBI" 
          } 
        }, 
        "assemblyNames": [ 
          "SL2.50" 
        ] } 
\end{lstlisting}

\subsection{Feature Search}

\subsection{Barcode Generation}

\section{User access of browser}

\end{document}

### 2025-04-28 Article/Article.tex
Article Draft ready for Tom

--- 2025-04-28 Article/Article.tex
@@
 \theoremstyle{thmstylethree}%
-\newtheorem{definition}{Definition}
+\newtheorem{definition}{Definition}%
+\usepackage{hyperref}%
 
@@
 
+% EARLIER TEXT KEEP UNTIL SAFE TO DELETE
 %\section{Expanding landscape of plant genomic studies}\label{sec2}
-
 % As previously shown by Kurowski & Mohareb (2020), Tersect, a lightweight, high-performing, real-time genome querying command-line utility that uses its own query language and fast execution through bitmap indices to interpret and apply flexible set theortical expressions to sets of sequence variant data, allows [something] identification of key variants and structures like haplotypes and introgressions in support of comparative genomic studies.
 Tersect returned
-
 % The field of tomato genomics has continued to evolve rapidly, with a growing emphasis on mapping introgressions — segments of DNA from wild tomato species that have been incorporated into cultivated lines.
@@
 \cite{RN7} , which sequenced 84 tomato accessions and wild relatives, exposed severe genetic bottlenecks in cultivated tomatoes and laid the groundwork for modern introgression analysis.
-
 % Recent studies have expanded on this base with increasingly sophisticated tools and datasets.
@@
 The construction of these pan-genomes enables researchers to observe variation absent in the reference genome, improving trait mapping and introgression detection.
-
 % \subsection{Background on introgression studies}\label{subsec1}
-
 % The role of introgression in water-tolerance Kubond et al. (2022), shows how such analysis will be important in identifying variant resistance to climate change.  
-
 % To manage and interpret the expanding complexity of genomic data, a wide range of bioinformatic tools have been deployed.
@@
 Visualization utilities such as the Introgression Browser (Aflitos et al., 2015) and the Tersect Browser (Kurowski and Mohareb, 2019) offer tailored environments for exploring introgression patterns, with Tersect’s set-theoretic framework enabling high-speed, complex genomic queries.
-
 % Meanwhile, long-read sequencing technologies (e.g., PacBio, Hi-C, and optical mapping) as applied in SL4.0 (Hosmani et al., 2019), and innovations in in situ hybridization (Shearer et al., 2014), have markedly improved the resolution of reference genomes—critical for detecting subtle and ancient introgressions.
-
 % Beyond tomato, research in soybean introgression offers valuable cross-crop insights.
@@
 Graph genome approaches and machine learning models are emerging as promising directions for future introgression prediction and trait association.
-
 % Together, these advances paint a dynamic picture of tomato genomics: one where introgression is not only a window into domestication and evolutionary history but also a powerful mechanism for ongoing improvement, including against threats such as climate change.
@@
 \section{Features}
-Tersect Browser+ is a web application that depends on resequenced data in VCF format and uses a reference genome in FASTA format.
-Utilizing a MEAN stack (MongoDB, Express, Angular, and Node) architecture, it creates a webpage displaying heatmaps and phylogenetic trees for introgressions within the loaded crop dataset.
-This provides an interface for users to identify differences across accessions within the same species dataset.
-It also includes a JBrowse interface for investigating specific variants, displaying selected genes directly alongside the generated heatmap.
-A search feature highlights high-impact variants on the webpage, helping to distinguish the selected reference genomes.
-Tersect Browser+ also offers barcode generation functionality for various cultivars and specific species in the dataset.
-
-\begin{figure} [http]
-  \centering
-  \includegraphics[width=\textwidth]{path/to_image.png}
-  \caption{Tersect Browser+ overview. The browser provides a single way to analyse variant data without leaving the page.[More description]}
-  \label{fig:your_label}
-\end{figure}
-
-\subsubsection{Gene Brwoser}
-
-We will include text here on the feature. I will look through the Technical Documentation to put this together.
-
-\subsubsection{Gene Model}
-
-We will include text here on the feature. I will look through the Technical Documentation to put this together.
-
-\subsubsection{Search Features}
-
-We will include text here on the feature. I will look through the Technical Documentation to put this together.
-
-\subsubsection{Barcode Generator}
-
-We will include text here on the feature. I will look through the Technical Documentation to put this together.
 
@@
 
+Tersect Browser+ is a web application that depends on resequenced data in VCF format and uses a reference genome in FASTA format.
+Utilizing a MEAN stack (MongoDB, Express, Angular, and Node) architecture, it creates a webpage displaying heatmaps and phylogenetic trees for introgressions within the loaded crop dataset.
+This provides an interface for users to identify differences across accessions within the same species dataset.
+It also includes a JBrowse interface for investigating specific variants, displaying selected genes directly alongside the generated heatmap.
+A search feature highlights high-impact variants on the webpage, helping to distinguish the selected reference genomes.
+Tersect Browser+ also offers barcode generation functionality for various cultivars and specific species in the dataset.
+
+\begin{figure} [http]
+  \centering
+  \includegraphics[width=0.5\textwidth]{path/to_image.png}
+  \caption{Tersect Browser+ overview. The browser provides a single way to analyse variant data without leaving the page.[More description]}
+  \label{fig:your_label}
+\end{figure}
+
+\subsubsection{Heatmap and Phylogenetic Tree}
+
+Visualising introgression distances for each accession against the reference genome provides cultivar biologists with insights into genetic similarity, divergence patterns, and potential introgressed regions that may influence desirable traits.
+Tersect Browser displayed a phylogenetic tree and corresponding heatmap for the selected reference genome, chromosome, chromosomal interval, and bin size.
+The heatmap represented the genetic distances between each accession and the reference genome (Kurowski, 2023).
+Tersect Browser+ extends and enhances these capabilities
+
+\subsubsection{Gene Browser and Model}
+
+Making it easier for users to explore gene models and genetic variants for a specific accession, this feature provides two ways to view the gene model track and variant track for the selected accession.
+The user can either select the ‘View Gene-Model’ button to load these tracks above the accession heatmap [Ref Technical Document Figure] or right-click on a heatmap region of interest to display the accession name and open the JBrowse interface [Add reference figure from the Technical document].
+This feature can also be used to [Add any other information on how this can be used].
+Gene model annotations are shown in the track [Add Technical Documentation figure reference].
+The paths are sorted and a compressed GFF file, along with its corresponding Tabix index are specified to the local server URLs during the browser start-up.
+
+\subsubsection{Search Features}
+
+High-impact variants, such as frameshift mutations, premature stop codons, or major missense mutations, can have profound effects on gene function.
+In crop breeding programs leveraging introgression, these variants are particularly important for assessing the functional consequences of transferred genomic regions.
+By enabling users to search for high-impact variants based on gene ID and impact severity, Tersect Browser+ facilitates the identification of critical genetic changes associated with desirable or undesirable traits.
+Selected variants are visually mapped onto the heatmap, highlighting the corresponding accession and genomic region [Figure from Technical Document reference].
+
+\subsubsection{Barcode Generator}
+
+Providing a DNA barcode or fingerprint of the genomic composition of each accession enables rapid identification of regions derived from donor versus recipient genomes.
+This allows breeders to verify the successful integration of target traits and to track the use of parental lines, which is useful in identifying specific accessions in the laboratory or published datasets.
+In recent years, significant advances have been made in DNA barcoding methodologies.
+For example, Prieto et al.
+(2024)\cite{espinosa_prieto_finding_2024} demonstrated its application in DNA fingerprinting of tomato breeding lines, while Al-Shammari et al.
+(2021) \cite{al2021genetic} documented its use in cultivar investigations.
+In Tersect Browser+, users can generate barcodes of their preferred length for any region on a selected chromosome.
+The tool identifies key characteristics of interest within the barcode region, including SNP counts and positions, the number of repeat elements, and GC content [Ref figure in technical documentation showing output].
+A default barcode length of 150 bp is provided, reflecting the attractiveness of lengths between 150–500 bp for primer production, as discussed by Letsiou et al.
+(2024) \cite{letsiou_dna_2024}.
+
+\subsubsection{Datasets}
+
+The Sol Genomics Network (SGN) tomato reference genome and associated resequenced accessions from cultivated lines and wild relatives was used.
+Reference genome updates (e.g., SL2.40 to SL2.50) were managed using SeqRemap, available under the MIT license at \url{https://bitbucket.org/cranfieldbix/seqremap}.
+
+\vspace{3mm}
+
+\noindent[Add on SoyBean]
+
 \section{Funding}
 
-A question for Tom - we see this included in many papers published in this journal. What should we include?
+[Tom - we see that funding sections are covered in many papers submitted to this journal. Do we need to include? If so, what do we need to include in here?]
 
-%\section{This is an example for first level head}\label{sec3}
+\vspace{3mm}
 
-% \subsection{This is an example for second level head - subsection head}\label{subsec2}
+\noindent\textit{Conflicts of interest}:~none declared.
 
-% \subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}
+\section{Supplementary Data}
 
-% \paragraph{This is an example for fourth level head - paragraph head}
+Technical documentation - pdf file
 
-% \section{Conclusion}
-
-% Some Conclusions here.
-
-%%%%%%%%%%%%%%
-
-\section{Code availability}
-
-We will add in a link to the github reference in here and a link to the technical documentation.
+\noindent\url{[githubreadme - to be added]}
 

### 2025-04-29 TechDoc/TechDoc.tex
Update TechDoc.tex
WIP: barcode technical description

@@
 \title{
-%{\includegraphics[width=3cm, height=2.5cm]{TechDoc/Cran.jpg}}
-%\\
-%\includegraphics[width=7cm, height=1cm]{TechDoc/TB.jpg}
-%\\
+{\includegraphics[width=3cm, height=2.5cm]{Cran.jpg}}
+\\
+\includegraphics[width=7cm, height=1cm]{TB.jpg}
+\\
 {Technical Documentation for TersectBrowser+}
@@
 This would allow for protection of intellectual property in the case of an advantageous new introgression that can be uniquely identified. 
-The Barcoding extension allows this unique identification by generating a range of barcodes with a user-specified maximum number of SNPs, so that primers for this barcode are easily generated in a wetlab environment.
-Metrics comparing these putative barcodes are also provided to the user, and available for local download.
+The Barcoding extension allows this unique accession-specific identification by generating a range of barcodes with a user-specified length and maximum number of SNPs,
+so that the barcodes are easily analysed in wetlab environments. Metrics comparing these putative barcodes are also provided to the user, and available for local download.
 
@@
 \paragraph{Feature Description}
-Given input parameters specified by the user, barcodes are generated via the backend and automatically downloaded for the user in txt format.
+Given input parameters specified by the user, barcodes are generated via the backend and can be downloaded in tsv format.
 The downloaded file is titled in the following format for easy identification and to prevent files being overwritten: \textit{SystemDateAndTime\_TB\_Barcode\_Gen\_AccessionName\.txt}.
@@
 \begin{itemize}
-    \item \textbf{Barcode sequence} - The base with the accession-specific SNP is enclosed in square brackets
-    \item \textbf{Barcode Start \& Barcode End} - The absolute position of the barcode in the chromosome 
-    \item \textbf{Variant Count} - The number of accession-specific SNPs that are present in the barcode 
-    \item \textbf{Variant Position} - The relative position of the accession-specific SNPs in the barcode 
-    \item \textbf{Repeat Sequence} - The sequence of the 2-bp repeating region. A repeating region is defined as 2 base pairs repeating 3+ times 
-    \item \textbf{Repeat Multiplier} - The number of times the repeat sequence is repeated 
-    \item \textbf{Repeat Start-End} - the relative start and end position of the repeat region within the barcode sequence 
-    \item \textbf{GC Content} - The GC content of the barcode 
+    \item \textbf{Barcode sequence} - The base with the accession-specific SNP is enclosed in square brackets.
+    \item \textbf{Chromosome} - The chromosome on which the barcode is located.
+    \item \textbf{Barcode Start \& Barcode End} - The absolute position of the barcode in the chromosome.
+    \item \textbf{Variant Count} - The number of accession-specific SNPs that are present in the barcode.
+    \item \textbf{Variant Position} - The relative position of the accession-specific SNPs in the barcode. 
+    \item \textbf{Repeat Sequence} - The sequence of the 2-bp repeating region. A repeating region is defined as 2 base pairs repeating 3+ times. 
+    \item \textbf{Repeat Multiplier} - The number of times the repeat sequence is repeated.
+    \item \textbf{Repeat Start-End} - the relative start and end position of the repeat region within the barcode sequence. 
+    \item \textbf{GC Content} - The GC content of the barcode. 
 \end{itemize}
@@
 Barcode generation is controlled by two scripts: \verb+barcode_finder.sh+ and \verb+find_barcode.py+.
+
+\paragraph{Bin menu barcode button}:
+A third menu item with the label \verb+Create barcode+ is added to the \verb+getAccessionItem()+ method in \verb+plot-click-menu.component+.
+The item is styled with the \verb+fa fa-barcode+ icon, and on command opens the \verb+ModalService openBarcodeModal()+ method, passing as arguments the plot state chromosome name and the bin accession label and start and end positions.
+
+\paragraph{Modal Service Component}:
 

### 2025-04-29 TechDoc/TechDoc.tex
Dataset setup instructions

@@
 This is where the server will look to find information on tracks when generating the Variant Browser panel.
+            \item This script checks for the install of modules such as SAMtools, BCFtools, and HTSlib, and will attempt to download these if not found.
+If a multi-sample VCF file is provided, it will split this file into individual sample level VCFs before proceeding.
+It will make sure these files are unzipped before using TERSECT CLI to build a tersect index for the whole dataset (it sources a previously created .tersect virtual environment to run this).
+After recompressing and indexing the fasta and VCF files, it will edit the \verb+add_example_dataset.sh+ script to point to the correct index file, and begin dataset addition to TersectBrowser+.
+Lastly, it will add the GFF file as a gene models track accessible within the browser.
         \end{itemize}
@@
 
-specified location within the repository: \textit{Tersect-browser/\(\sim \)/mongo-data/gp\_data\_copy/}
-\begin{itemize}
-    \item Download all VCF files in the resequenced dataset. Run tabix on these files to create .tbi index files.
-    \item Download the reference genome FASTA file. Run SAMtools faidx to create .fai index file.
-    \item Download (or generate) an annotation GFF file for the resequenced dataset using SNPeff.  Index the GFF file. Run tabix on this file to create a .tbi index file.
-\end{itemize}
-\begin{itemize}
-\item Use JBrowse CLI text-index tool to generate Trix text searching files (.ix, .ixx, and .meta.json) from the GFF annotation file.
-\item Run the script \verb+add_config_tracks.py+ using these downloaded files. This will use a set of jbrowse arguments to load the datafiles as tracks in the Variant Browser component.
-\item Run the script \verb+change_dataset_files.py+ using these downloaded files as CLI arguments, in order to update the references to these data files in the project.
-\item Run the script \verb+add_example_dataset.sh+ to prepare the dataset.
-\end{itemize}
+\subsection{Deployment with loaded dataset}
+\begin{enumerate}
+    \item Run the Mongo Database using:
+\end{enumerate}
 

### 2025-04-29 TechDoc/TechDoc.tex
Deployment information

@@
 \begin{enumerate}
-    \item Run the Mongo Database using:
+    \item Run the Mongo Database using: \\
+        \verb+mongod --dbpath <pathto/mongo-data>+
+    \item Change to the root of the Tersect-browser repo and run:\\
+        \verb+nvm use 16+\\
+        \verb+npm start+
+    \item Change to the ersect-browser/extension/genome-browser/ directory and run:\\
+        \verb+nvm use 22+\\
+        \verb+npm start+
+    \item Open TersectBrowser+ on \url{http://localhost:4200/TersectBrowserGP/}
 \end{enumerate}
@@
 \section{TersectBrowser+}
+The project brief to extend the original TersectBrowser listed four main extension requirements, as well as a stretch extension.
+These requirements were provided as discrete goals for the project, and informed the main working priorities for the project timeline, however in practice the four extensions do have significant overlap in terms of code structure.
+For this reason, the accompanying TersectBrowser+ Article includes the extensions Variant Browser and Gene Models in the same section, while this Technical Documentation keeps the two separate according to the chronological timeline of developments to these extensions.
+Architecture diagrams are used to display the overlap between various extension components, especially relating to how they orchestrate data flow through the entire software.
+
 \subsection{Variant Browser Extension}

### 2025-04-29 TechDoc/TechDoc.tex

@@
 Architecture diagrams are used to display the overlap between various extension components, especially relating to how they orchestrate data flow through the entire software.
+\\
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=15cm, height=9cm]{high_level_arch_draft.png}
+    \caption{Summary Architecture of TersectBrowser+.}
+\end{figure}
 
@@
 \subsubsection{User Interface}
-\paragraph{Feature Description}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=5cm, height=9cm]{UserBarcode.jpg}
+    \caption{User-focused data flow through Barcoding Extension.}
+\end{figure}
+\paragraph{Frontend Display Feature Description}
+When the user clicks on a bin, the popup menu appears and the button "Create barcode" is visible, styled with a barcode icon.
+When the user clicks on this button, a popup window appears in the centre of the screen.
+The window title is the corresponding accession name.
+The window contains the header "Barcode Generator" and contains a description of what is being generated.
+There is a box for the user to input barcode size, with default set to 150, and the maximum number of variants allowed per barcode (this is an optional parameter.
+Leaving this blank means all possible barcodes will be returned).
+
+At the bottom is a green "Generate" button.
+Hovering over the button turns it dark green, and clicking on this button will call the backend scripts to generate the barcodes output.
+Whilst scripts are running and barcodes are being generated, a spinner is added to the button to indicate loading.
+After the script has been generated, the spinner disappears and another green button "Download" appears at the bottom left of the window.
+Clicking on the button downloads the tsv file containing the barcodes.
+
+\paragraph{Backend Script Feature Description}
 Given input parameters specified by the user, barcodes are generated via the backend and can be downloaded in tsv format.
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=7cm, height=12cm]{ExtBarcode.jpg}
+    \caption{User-focused data flow through Barcoding Extension.}
+\end{figure}
+
 Barcode generation is controlled by two scripts: \verb+barcode_finder.sh+ and \verb+find_barcode.py+.
 
-\paragraph{Bin menu barcode button}:
+\paragraph{Bin menu barcode button}:\\
 A third menu item with the label \verb+Create barcode+ is added to the \verb+getAccessionItem()+ method in \verb+plot-click-menu.component+.
@@
 
-\paragraph{Modal Service Component}:
+\paragraph{Modal Service Component}:\\
+The observables \verb+barcodeVisible$+, \verb+barcodeTitle$+, \verb+barcodeChromo$+, \verb+barcodeStart$+ and \verb+barcodeEnd$+ are added to \verb+modal.service+.
+The method \verb+openBarcodeModal()+ is added, accepting as arguments \verb+accessionName+, \verb+chrom+, \verb+startPosition+, and \verb+endPosition+.
+These arguments are saved to their corresponding observables.
 
-\paragraph{Calling Tersect CLI to extract variants}: 
+A new modal, \verb+global-barcode-modal+, is created.
+The modal is imported to app.module and added to app.component.
+The \verb+global-barcode.component.ts+ imports \verb+ModalService+ and \verb+TersectBackendService+.
+Variables to hold the modal state, modal loading state, modal title, barcode size, chromosome, interval start position, interval end position, maximum variants per barcode, the downloadable URL and filename are defined.
+On initialisation, modal state, modal title, chromosome, interval start position and interval end position are saved to observables in modal.service.
+
+The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to "True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+.
+The returned URL and filename are saved to the variables \verb+downloadUrl+ and \verb+downloadFileName+.
+Once returned, \verb+isLoading+ is set to "False" and \verb+barcodeSize+ is set back to "150".
+
+The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to "150", \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the "Download" button from the popup.
+
+\paragraph{Modal Service Styling}:\\
+The modal template is defined in \verb+global-barcode.component.html+.
+The popup window is visible depending on the modal state isVisible, and is centrally position with \verb+modalTitle+ set as the header.
+Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button "Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated.
+An anchor displays a clickable "Download" p-button that points towards the generated barcode file.
+The filename is set to \verb+downloadFileName+, and "Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.
+
+Styling is defined in the \verb+global-barcode.component.css+.
+Button background colour is set to \verb+#459e00+, background hover colour is set to \verb+#33b357+, and text colour is set to white.
+The input boxes and labels are given a set width of 200px.
+
+\paragraph{Router and Backend service}:\\
+A new route, \verb+/generate-barcodes+, is added to \verb+tersect-router+.
+This router receives in the request the following parameters: \verb+accessionName+, \verb+chrom+, \verb+start+, \verb+end+, \verb+size+ and \verb+maxVar+.
+Paths to the tsi index and fasta file are defined.
+The backend script, \verb+barcode_finder.sh+, is called, with the above parameters passed as arguments.
+A \verb+downloadableURL+ is created, pointing to the created file, and returned in json format.
+
+The method \verb+generateBarcodes()+, is added to \verb+tersect-backend.service+.
+It sends a HTTP Post request to the endpoint \verb+/generate-barcodes+, triggering barcode generation.
+The method returns an observable containing the \verb+downloadableURL+ pointing to the generated barcode file.
+The method passes the following parameters:
+\begin{itemize}
+    \item \textbf{accessionName} - a string containing the specified accession name for which to generate unique barcodes 
+    \item \textbf{Chrom} - a string containing the chromosome identifier for the chromosome where the barcodes will be generated 
+    \item \textbf{Start} - a number specifying the start position of the interval in which to generate barcodes 
+    \item \textbf{End} - a number specifying the end position of the interval in which to generate barcodes 
+    \item \textbf{Size} - a number specifying the length of the barcodes 
+    \item \textbf{maxVar} - an optional parameter, specifying the maximum permissible number of variants located within the barcode.
+Either a number or null, indicating all barcodes with all numbers of variants are generated
+\end{itemize}
+
+\paragraph{Calling Tersect CLI to extract variants}: \\
 \verb+Barcode_finder.sh+ is run on the command line, and takes as input accession name, chromosome, interval start position, interval end position, barcode size, maximum variant number, reference fasta, and tersect TSI index.
@@
 
-\paragraph{Generating barcodes}: 
+\paragraph{Generating barcodes}: \\
 \verb+Find_barcode.py+ requires the following dependencies: argparse, SeqIO, and datetime.
@@
 
-\paragraph{Output file and barcode stats}: 
+\paragraph{Output file and barcode stats}: \\
 Statistical metrics are calculated for each barcode, using custom methods.
@@
 \subsubsection{Automated Introgression Search Extension}
-\section{References}
-
-\subsection{Purpose}
-Identify the purpose of this SDD and its intended audience. (e.g. This software design
-document describes the architecture and system design of XX.. ).
-
-\subsection{Scope}
-Provide a description and scope of the software and explain the goals, objectives and benefits
-of your project. This will provide the basis for the brief description of your product.
-
-\subsection{Overview}
-Provide an overview of this document and its organization.
-
-\subsection{Reference Material}
-This section is optional.
-List any documents, if any, which were used as sources of information for the test plan.
-
-\subsection{Definitions and Acronyms}
-This section is optional.
-Provide  definitions of all terms, acronyms, and  abbreviations that might exist to  properly 
-interpret the SDD. These definitions should be items used in the SDD that are most likely not 
-known to the audience.
-
-\begin{table}[h]
-\begin{tabular}{|l|l|}
-\hline
-\multicolumn{1}{|c|}{\textbf{Term}} & \multicolumn{1}{c|}{\textbf{Definition}}                                                                                              \\ \hline
-Software Design Document (SDD)      & \begin{tabular}[c]{@{}l@{}}Used as the primary medium for communicating software \\ design information.\end{tabular}                  \\ \hline
-Design Entity                       & \begin{tabular}[c]{@{}l@{}}An element of a design that is structurally and functionally \\ distinct from other elements.\end{tabular} \\ \hline
-\end{tabular}
-\end{table}
-
-\section{System Architecture}
-\subsection{Architectural Design}
-
-\subsection{Design Rationale}
-
-\subsection{Data Description}
-
-\subsection {Data Dictionary}
 

### 2025-04-30 TechDoc/TechDoc.tex
Added contents page
Added FeatureSearch clear button description

@@
 \paragraph{\hl{Clear Button }}
+The clear button is located to the right of the \verb+Open variant search+ button and styled with \verb+fa fa-refresh+.
+It contains the event handler \verb+(click)="refreshBin()"+ that calls the \verb+refreshBin()+ method in \verb+BinDrawService+.
+This method calls \verb+generatePlotArray()+, which re-creates the bin view, and \verb+updateCanvas()+, which redraws the canvas using the re-created bin view.
 

### 2025-04-30 TechDoc/TechDoc.tex
Edits to genome browser feature description

@@
 
-\subsection{Variant Browser Extension}
+\subsection{Genome Browser Extension}
 \label{sec:Browser}
@@
 However, in order for TersectBrowser+ to be competitive as a tool with other browsers, it was necessary to implement features that enhance further investigation within the scope of the same browser.
-The main view of the original TersectBrowser displayed the density of variants along each chromosome of the reference genome, a view compatible with the addition of a variant browser with multiple tracks providing extra information at a per-base level.
+The main view of the original TersectBrowser displayed the density of variants along each chromosome in comparison to the reference genome.
+To enhance this view, a variant browser with multiple tracks providing extra information at a per-base level was added to enable to user to directly correlate variant density with accession-specific variants.
 
-After a review of node-compatible browser tools such as CRAMER, JBrowse was ultimately chosen to carry out this extension request.
+After a review of node-compatible browser tools such as CRAMER, JBrowse was ultimately chosen to carry out this extension request, due to its high customisability, up-to-date features, and active technical support.
 The initial build of JBrowse into TersectBrowser+ as an extension focused on retaining features most relevant to the user aims, and bringing the styling of the panel in line with the rest of the page.
@@
 \begin{itemize}
-    \item Window is located at top of Tersect Browser (TB)- below settings bar but above TB heatmap 
-    \item Jbrowse window (JB) layout - shows only sequence tracks and ruler track/linear genome scale.
-Zoom features/chromosome selection features have been removed from Jbrowse window.
-Hamburger button opens menu containing many options, notably track selector, where tracks can be added/removed.
-The chromosome scale for TB has a dynamic offset depending on the accession/tree view and the zoom level.
-This is synced between JB and TB, so that the chromosome scale lines up.
+    \item JBrowse window:
+    \begin{itemize}
+        \item Main window is located at the top of Tersect Browser, below settings bar but above the heatmap 
+        \item Main window shows only sequence tracks and linear genome scale.
+Interactive features such as zoom anc chromosome selection have been removed from the main window.
+The container window has a dynamic offset, which changes according to tree view and zoom level to ensure the chromosome scales between TersectBrowser and JBrowse line up.
+        \item By default, the main window shows only the gene models track.
+        \item Popup window provides a more sandbox experience, providing more user interaction.
+The user can change zoom level and displayed chromosome, and two scale bars enable both precise and rapid navigation.
+The hamburger menu on the top left provides the capability for the user to upload their own data to visualise unique tracks.
+        \item By default, the popup window shows the reference sequence track, gene models track, and variant track for the specified accession.
     \begin{itemize}
@@
     \end{itemize}
-    \item Variant browser window is preloaded with default tracks. 
+    \item Syncing of the JBrowse window and the TersectBrowser heatmap includes the following:
     \begin{itemize}
-        \item \hl{Track TBC}: Currently pre-loaded with first 3 tracks from dataset. Should be changed to 1-2 meaningful accessions?
-    \end{itemize}
-    \item Syncing of the Browser pane and the Tersect heatmap pane, including the following:
-    \begin{itemize}
-        \item Zoom synced, both when the zoom plus/minus buttons are pressed, and also when scrolling the mouse to zoom. 
-        \item Bin size changes synced.
-        \item Interval synced using an offset of the Browser pane.
-        \item Chromosome selection changes synced.
-        \item Horizontal scrolling along the Browser and Tersect panes is synced.
+        \item Zoom - both by clicking the plus/minus buttons and by scrolling with the mouse wheel.
+        \item Bin size.
+        \item Selected interval.
+        \item Chromosome.
+        \item Horizontal scrolling along the x-axis.
     \end{itemize}

### 2025-04-30 TechDoc/TechDoc.tex

@@
 \subsubsection{Rationale}
-Once the Variant Browser panel was added to the main page of TersectBrowser+, this facilitated the inclusion of a specified gene model track.
-This allows the plant breeder user to identify an area of high variant density on the heatmap, and then check the gene model track to investigate whether this region occurs near a genetic feature with functional relevance to the species.
+Once the Genome Browser panel was added to the main page of TersectBrowser+, this facilitated the inclusion of a reference gene model track.
+This allows the plant breeder user to identify an area of high variant density on the heatmap, and then check the gene model track to investigate whether this region occurs near a gene of interest.
 
@@
 \begin{itemize}
-    \item In the Jbrowse window, the first track shows gene models based on the GFF file provided.  
+    \item The gene models track is displayed in the main JBrowse window. 
 %    \item Gene model annotations are based on the official \hl{tomato genome annotation v2.4}: \href{https://phytozome-next.jgi.doe.gov/info/Slycopersicum_ITAG2_4#:~:text=Genome%20Overview,genome%2Cwith%2034%2C725%20gene%20models}{Slycopersicum ITAG annotation v2.4}
-% The above is tomato-specific. Not sure where it fits
-    \item The popup view of the same Jbrowse component has similar default: the first track is the gene models track, and the second track is the variant track for the selected accession.
+% The above is tomato-specific. Not sure where it fits. Probably doesn't fit anywhere - included it in notes so that we knew where the tomato gene models came from
+    \item In the popup window, the gene models track is displayed as the second track, underneath the reference sequence track and above the variant track.
 \end{itemize}
@@
 The Jbrowse \verb+viewState+ is configured in \verb+JbrowseWrapper.tsx+, as described in Extension-JBrowse.
-If no accession is selected, the first three tracks stored in \verb+tracks.ts+ are added to the \verb+viewState+.
-As the Gene Models \verb+FeatureTrack+ is the first track in \verb+tracks.ts+, it is displayed as the first track in the Jbrowse window.
+If no accession is selected, the first track stored in \verb+tracks.ts+ (gene models track) is added to the \verb+viewState+.
 
 For the popup window, the JBrowse \verb+viewState+ is configured when an accession is selected in \verb+JbrowseWithAccession.tsx+, as described in \nameref{sec:Browser}.
-If a track with a \verb+trackId+ matching the selected accession is found, the viewer displays both the first track defined in \verb+tracks.ts+ and the matching track.
+If a track with a \verb+trackId+ matching the selected accession is found, the viewer displays the reference sequence track from \verb+assembly.ts+, the first track defined in \verb+tracks.ts+ (gene model track) and the selected accession variant track.
 
@@
 \subsubsection{Rationale}
-The aim of this extension is to allow the user to search specific genes and identify which accession contains a \hl{high/medium/low} impact variant affecting that gene.
+The aim of this extension is to allow the user to search selected intervals and identify which accessions contain \hl{high/medium/low} impact variants within that region.
 The bins containing high impact variants for that gene will be highlighted at the accession level, allowing high resolution investigation of variants.
-The user may alternatively search for any high impact variants within a specified region, in case they lack a specific gene name to search.
 The highlighted bins may then be selected and opened in the popup \nameref{sec:Browser} window for further investigation.
@@
 \begin{itemize}
-    \item Separate search bar in the top right of the tersect browser header where user can input gene name and search button 
-    \item \hl{TBC: Option} for user to select based on region in addition to gene name 
-    \item Popup window for user to select advanced features 
-    \item Output: Bins in the canvas are highlighted red at the chromosomal position where the gene is located, and only for accessions containing a variant impacting that gene 
-    \item Clear button to clear highlighted bins from canvas 
+    \item \verb+Open Variant Search+ button in TersectBrowser+ header, opens popup window for user to select advanced features.
+    \item Popup window where user can select interval, choose impact level, and search button to initiate search.
+    \item Output: Bins in the canvas are highlighted red at the chromosomal position where the variant is located, and only for accessions containing variants of the specified impact.
+    \item Clear button located in TersectBrowser+ header to clear highlighted bins from canvas and restore original heatmmap.
 \end{itemize}

### 2025-05-02 TechDoc/TechDoc.tex
Small edits to intro

@@
 \section{Overview}
-This Technical Documentation describes the particular details of updating the original TersectBrowser developed by Tomasz Kurowski into the TersectBrowser+ by integrating multiple extension features.
+This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
+It outlines the integration of multiple extension features, providing the user with additional display and data analysis capabilities.
 
@@
 The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
-It was chosen to use a 'bolt-on' approach for the extensions in order to fit the project delivery within the time requirements, as well as allowing the new extensions to use up-to-date packages.
-The app is dockerized to allow both versions of node to be used seamlessly in different sections.
+This 'bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
+Additionally, this approach allows the new extnesions to use up-to-date packages, resulting in enhanced functionality and a modern design.
+To run TersectBrowser+ on different platforms, the new extensions are hosted on NPM and can be linked as dependencies within the project.
 
-For the TersectBrowser+ update to the original TersectBrowser, main changes have been made to the frontend implementation of features added in a separate extension section of the project.
-The main addition is in a 'genome-browser' extension, which contains elements for each of the specific features mentioned in the project brief.
+Extensions added to TersectBrowser+ include the integration of a genome browser capable of displaying a variety of different tracks, notably variants and gene models, search functionality to identify and display regions containing high impact variants, and a tool to generate accession-specific barcodes.
+These extensions are hosted in a separate section of the project, but interact with both frontent and backend services to provide a seamless user experience.
 
 \section{Deployment}
-The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment, however the main outline is given here.
+The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment.
+Here, we expand upon these instructions and give a detailed overview of the structure of each extension.
 
@@
 These requirements were provided as discrete goals for the project, and informed the main working priorities for the project timeline, however in practice the four extensions do have significant overlap in terms of code structure.
-For this reason, the accompanying TersectBrowser+ Article includes the extensions Variant Browser and Gene Models in the same section, while this Technical Documentation keeps the two separate according to the chronological timeline of developments to these extensions.
+For this reason, the accompanying TersectBrowser+ Article includes the extensions Genome Browser and Gene Models in the same section, while this Technical Documentation keeps the two separate according to the chronological timeline of developments to these extensions.
 Architecture diagrams are used to display the overlap between various extension components, especially relating to how they orchestrate data flow through the entire software.
@@
         \item By default, the popup window shows the reference sequence track, gene models track, and variant track for the specified accession.
-    \begin{itemize}
-        \item \hl{Layout TBC}: Set window size
-    \end{itemize}
     \item Tracks:
@@
 This would allow for protection of intellectual property in the case of an advantageous new introgression that can be uniquely identified.
-The Barcoding extension allows this unique accession-specific identification by generating a range of barcodes with a user-specified length and maximum number of SNPs,
-so that the barcodes are easily analysed in wetlab environments. Metrics comparing these putative barcodes are also provided to the user, and available for local download.
+The Barcoding extension allows this unique accession-specific identification by generating a range of barcodes with a user-specified length and maximum number of SNPs, so that the barcodes are easily analysed in wetlab environments.
+Metrics comparing these putative barcodes are also provided to the user, and available for local download.
 

### 2025-05-02 TechDoc/TechDoc.tex
Edits to genome browser & gene models section

@@
 
-The function \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+.
-If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the constant \verb +JbrowseWithAccessionName+ is returned (See below parameters).
+The method \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+.
+If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the \verb +JbrowseWithAccessionName+ is returned (See below parameters).
 If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated.
@@
 \end{itemize}
-
-When no accession is selected in the Canvas interface, the first three tracks defined in \verb +tracks.ts+ are added to the \verb +viewState+.
-The dynamic left-hand offset is set by \verb +horizontalScroll()+.
 
@@
 \paragraph{Offset}
-The dynamic offset is synced in a similar fashion to zoom, bin size, chromosome selection, and interval.
+The left-margin offset requires dynamic syncing to ensure the scale between JBrowse and the heatmap remain aligned.
 The observable \verb +offsetCanvas+ is defined as a component of the \verb +PlotStateService+ class.
@@
 
-Inside \verb +tersect-browser.component.ts+, the variable \verb +offsetCanvas+ is defined, and the subscription \verb +offsetCanvasSub+ listens to the  \verb +offsetCanvas+ observable and assigns the latest value to \verb +offsetCanvas+.
-Inside \verb +tersect-browser.component.html+, the \verb +offsetCanvas+ is passed to \verb +JbrowseWrapper+ as the prop \verb +offsetCanvas+.
-Inside \verb +JbrowseWrapper+, the variable defines the extent of the \verb +horizontalScroll+ using the following logic: $horizontalScroll(-(location.offsetCanvas - 4))$.
+The position of the JBrowse container is dynamically set in \verb+tersect-browser.component.ts+, with the \verb+style.margin-left.px+ set to \verb+offsetCanvas+, which holds thw pixel width of the phenetic tree.
+As a result, the position of the JBrowse container dynamically adjusts to different canvas states to ensure the scale bar between JBrowse and the heatmap remains aligned.
+
+\paragraph{Horizontal Scroll}
+Horizontal scroll inputs from the user are disabled inside the main JBrowse window through the implementation of a \verb+Plugin+ of the \verb+DisableScroll+ class in \verb+JbrowserWrapper.tsx+, which overrides the JBrowse configuration to prevent horizontal scrolling
 
@@
 When a new dataset is added, the custom addition script automatically updates the names of tracks and assembly in the relevant files within the genome-browser extension.
-This flexibility allows the Variant Browser extension to be accessible by plant breeders working on many different datasets.
+This flexibility allows the Genome Browser extension to be accessible by plant breeders working on many different datasets.
 TersectBrowser was initially validated with a human genome dataset, so it was important to maintain this flexibility of input in the new extensions.
@@
 \subsubsection{Future Improvements}
+In addition to displaying the reference track, gene models track, and variants track, additional feature tracks, such as a multiple sequence alignment view, could be added to the genome browser.
+
+Currently, the genome browser popup window is limited to opening a single window within the TersectBrowser+ webpage.
+A button to open the popup window in a new tab could be added, which would allow the user to open multiple detatched popup windows and view them simultaneously.
 
@@
     \item The gene models track is displayed in the main JBrowse window. 
-%    \item Gene model annotations are based on the official \hl{tomato genome annotation v2.4}: \href{https://phytozome-next.jgi.doe.gov/info/Slycopersicum_ITAG2_4#:~:text=Genome%20Overview,genome%2Cwith%2034%2C725%20gene%20models}{Slycopersicum ITAG annotation v2.4}
-% The above is tomato-specific. Not sure where it fits. Probably doesn't fit anywhere - included it in notes so that we knew where the tomato gene models came from
     \item In the popup window, the gene models track is displayed as the second track, underneath the reference sequence track and above the variant track.

### 2025-05-02 TechDoc/TechDoc.tex
added to future extension work

@@
 \subsubsection{Test Results}
+
 \subsubsection{Future Improvements}
+Currently, only bins on the TersectBrowser+ heatmap are highlighted in red to indicate the positions of high-impact variants.
+Highlighting could be extended across the Genome Browser and Gene Models extension, so that gene models matching the parameters of the feature search query are also highlighted in red.
 
@@
 \subsubsection{Future Improvements}
+Currently, barcode viewing is limited to the downloaded file. It can be extended to display barcodes in the popup window, with the variant highlighted in red.
+
 \subsection{Future Work and Extension Design}

### 2025-05-05 TechDoc/TechDoc.tex
Adding in diagrams

@@
 \subsubsection{User Interface}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=10cm, height=8cm]{GenomeBrowser.png}
+    \caption{UI diagram for Genome Browser.}
+\end{figure}
 
@@
         \item Horizontal scrolling along the x-axis.
+    \end{itemize}
 \end{itemize}
@@
 \subsubsection{User Interface}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=15cm, height=9cm]{FS1.png}
+    \caption{UI digram for Feature Search.}
+\end{figure}
+
 \paragraph{Feature Description}
@@
 \subsubsection{Technical Description}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=7cm, height=15cm]{FS2.png}
+    \caption{Data flow diagram for Feature Search.}
+\end{figure}
+
 \paragraph{\hl{Search Bar}}
@@
     \centering
-    \includegraphics[width=5cm, height=9cm]{UserBarcode.jpg}
-    \caption{User-focused data flow through Barcoding Extension.}
+    \includegraphics[width=6cm, height=8cm]{Bar1.png}
+    \caption{UI diagram for Barcode Generation.}
 \end{figure}
+
 \paragraph{Frontend Display Feature Description}
@@
     \centering
-    \includegraphics[width=7cm, height=12cm]{ExtBarcode.jpg}
-    \caption{User-focused data flow through Barcoding Extension.}
+    \includegraphics[width=7cm, height=15cm]{Bar2.png}
+    \caption{Data flow diagram for Barcode Generation.}
 \end{figure}

### 2025-05-05 TechDoc/TechDoc.tex
Formatting

@@
 The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
-This 'bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
+This `bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
 Additionally, this approach allows the new extnesions to use up-to-date packages, resulting in enhanced functionality and a modern design.
@@
     \item A GFF file of gene model information for the reference genome, produced using SNPeff software.
-If not using SNPeff, ensure that variant impact levels are categorised into 'High', 'Medium', and 'Low' impact.
+If not using SNPeff, ensure that variant impact levels are categorised into `High', `Medium', and `Low' impact.
 If not provided, the gene model track in the Variant Browser will be unavailable.
     \item Whether there are prior identified introgressions known to be functionally relevant for the species.
-These can be used to 'check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
+These can be used to `check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
 \end{itemize}
@@
 If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated.
-If either of them are empty, a container with the phrase "Loading…" is returned.
+If either of them are empty, a container with the phrase ``Loading…" is returned.
 If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+.
@@
 The configurations for the assembly track, which is built upon the reference accession SL2.50, is defined in \verb +assembly.ts+ in json format.
-The name is set to the reference accession name and \verb +trackId+ is set to "SL2.50-ReferenceSequenceTrack".
+The name is set to the reference accession name and \verb +trackId+ is set to ``SL2.50-ReferenceSequenceTrack".
 The paths to the fasta file, along with its corresponding fasta index, are specified as local server URLs provided by the backend during the Tersect Browser setup.
@@
 The configurations for the \verb +VariantTrack+ and Gene Models \verb +FeatureTrack+ are defined in \verb +tracks.ts+ in json format.
-For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to "ITAG2.4 Gene Models".
+For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to ``ITAG2.4 Gene Models".
 The paths to the zipped files, along with its corresponding Tabix index files, are specified as local server URLs provided by the backend during the Tersect Browser setup.
@@
 \paragraph{Styling}
-The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour '\#459e00' and the \verb +boxShadow+ is set to 'none'.
+The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour `\#459e00' and the \verb +boxShadow+ is set to `none'.
 
@@
 \paragraph{Frontend Display Feature Description}
-When the user clicks on a bin, the popup menu appears and the button "Create barcode" is visible, styled with a barcode icon.
+When the user clicks on a bin, the popup menu appears and the button ``Create barcode" is visible, styled with a barcode icon.
 When the user clicks on this button, a popup window appears in the centre of the screen.
 The window title is the corresponding accession name.
-The window contains the header "Barcode Generator" and contains a description of what is being generated.
+The window contains the header ``Barcode Generator" and contains a description of what is being generated.
 There is a box for the user to input barcode size, with default set to 150, and the maximum number of variants allowed per barcode (this is an optional parameter.
@@
 
-At the bottom is a green "Generate" button.
+At the bottom is a green ``Generate" button.
 Hovering over the button turns it dark green, and clicking on this button will call the backend scripts to generate the barcodes output.
 Whilst scripts are running and barcodes are being generated, a spinner is added to the button to indicate loading.
-After the script has been generated, the spinner disappears and another green button "Download" appears at the bottom left of the window.
+After the script has been generated, the spinner disappears and another green button ``Download" appears at the bottom left of the window.
 Clicking on the button downloads the tsv file containing the barcodes.
@@
 
-The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to "True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+.
+The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to ``True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+.
 The returned URL and filename are saved to the variables \verb+downloadUrl+ and \verb+downloadFileName+.
-Once returned, \verb+isLoading+ is set to "False" and \verb+barcodeSize+ is set back to "150".
+Once returned, \verb+isLoading+ is set to ``False" and \verb+barcodeSize+ is set back to ``150".
 
-The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to "150", \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the "Download" button from the popup.
+The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to ``150``, \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the ``Download" button from the popup.
 
@@
 The popup window is visible depending on the modal state isVisible, and is centrally position with \verb+modalTitle+ set as the header.
-Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button "Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated.
-An anchor displays a clickable "Download" p-button that points towards the generated barcode file.
-The filename is set to \verb+downloadFileName+, and "Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.
+Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button ``Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated.
+An anchor displays a clickable ``Download" p-button that points towards the generated barcode file.
+The filename is set to \verb+downloadFileName+, and ``Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.
 
@@
 
-The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+'TB_Barcode_Gen'+, and the specified accession.
+The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+`TB_Barcode_Gen'+, and the specified accession.
 

### 2025-05-05 TechDoc/TechDoc.tex
Updates from Overleaf

--- 2025-05-02 TechDoc/TechDoc.tex
@@
 \section{Overview}
-This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
+This document describes the technical specifications and extended functionalities of Tersect Browser+, a web-based application for the exploration of genome and variant data in cultivated and wild-type crops.
+Building upon the original Tersect Browser, developed by Tomasz Kurowski [Reference], this enhanced version integrates additional features to support the identification and analysis of introgressions within accession species.
+The system leverages a MongoDB database to manage genomic datasets efficiently and provides a suite of interactive tools for biological interpretation.
+
+%This document describes the technical specifications particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
 It outlines the integration of multiple extension features, providing the user with additional display and data analysis capabilities.
 
+The original Tersect Browser offered a streamlined interface for investigating genomic intervals and phylogenetic relationships, powered by the high-performance Tersect CLI.
+It included dynamic visualizations such as phylogenetic trees and genetic distance heatmaps derived from resequenced genome data.
+
+Tersect Browser+ extends this foundation with several key enhancements aimed at cultivar biology.
+These include a genome browser, gene model visualizer, feature search tools, and barcode generation utilities—together offering a more powerful and versatile platform for crop genetics research.
+
 \section{Architecture}
-The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
+The Tersect Browser+ application is composed of four main components: 
+\begin{itemize}
+    \item the frontend, 
+    \item backend, 
+    \item common utilitises, and
+    \item extensions modules
+\end{itemize}
+
+The frontend is a browser-based interface developed in Angular and React, designed for interactive data visualization, including distance plots, accession selectors, and genome browser integration.
+The backend, built in Node.js and Express, provides API endpoints, data models, and utility scripts to serve genomic data and computation results from a MongoDB database.
+A set of common modules supports both frontend and backend logic with shared resources.
+Additionally, the architecture supports pluggable extensions, such as the genome-browser, which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons.
+Supporting data files, environment configurations, and Python scripting for preprocessing or bioinformatics workflows are organized in separate folders for virtual environments and data loading, enabling efficient pipeline integration and deployment.
+
+%The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
 This 'bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
@@
 
+Tersect Browser+ is primarily built using Angular 8, preserving the architecture of the original application.
+To enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22.
+This modular, "bolt-on" extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend and backend technologies.
+Extensions are developed independently and published to NPM, making them easily linkable through project dependencies.
+This hybrid approach ensures forward compatibility and scalability, while maintaining stability in the core application.
+Future versions of Tersect Browser+ may explore fully migrating the main application to a more unified and modern framework.
+
+\subsection{Use Cases}
+
 Extensions added to TersectBrowser+ include the integration of a genome browser capable of displaying a variety of different tracks, notably variants and gene models, search functionality to identify and display regions containing high impact variants, and a tool to generate accession-specific barcodes.
 These extensions are hosted in a separate section of the project, but interact with both frontent and backend services to provide a seamless user experience.
+
+\subsection{High Level Architecture}
 
@@
 
-\subsection{Dataset Upload}
+\subsection{Databases}
+
+\begin{enumerate}
+    \item Run the Mongo Database using: \\
+        \verb+mongod --dbpath <pathto/mongo-data>+
+    \item Change to the root of the Tersect-browser repo and run:\\
+        \verb+nvm use 16+\\
+        \verb+npm start+
+    \item Change to the ersect-browser/extension/genome-browser/ directory and run:\\
+        \verb+nvm use 22+\\
+        \verb+npm start+
+    \item Open TersectBrowser+ on \url{http://localhost:4200/TersectBrowserGP/}
+\end{enumerate}
+
+\subsection{Loaded Datasets}
 
@@
         \end{itemize}
-\end{enumerate}
-
-\subsection{Deployment with loaded dataset}
-\begin{enumerate}
-    \item Run the Mongo Database using: \\
-        \verb+mongod --dbpath <pathto/mongo-data>+
-    \item Change to the root of the Tersect-browser repo and run:\\
-        \verb+nvm use 16+\\
-        \verb+npm start+
-    \item Change to the ersect-browser/extension/genome-browser/ directory and run:\\
-        \verb+nvm use 22+\\
-        \verb+npm start+
-    \item Open TersectBrowser+ on \url{http://localhost:4200/TersectBrowserGP/}
 \end{enumerate}

### 2025-05-05 TechDoc/TechDoc.tex
Some changes to the overview section and architecture opening.

--- 2025-05-05 TechDoc/TechDoc.tex
@@
 \section{Overview}
-This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
+This document describes the technical specifications and extended functionalities of Tersect Browser+, a web-based application for the exploration of genome and variant data in cultivated and wild-type crops.
+Building upon the original Tersect Browser, developed by Tomasz Kurowski [Reference], this enhanced version integrates additional features to support the identification and analysis of introgressions within accession species.
+The system leverages a MongoDB database to manage genomic datasets efficiently and provides a suite of interactive tools for biological interpretation.
+
+The original Tersect Browser offered a streamlined interface for investigating genomic intervals and phylogenetic relationships, powered by the high-performance Tersect CLI.
+It included dynamic visualizations such as phylogenetic trees and genetic distance heatmaps derived from resequenced genome data.
+
+Tersect Browser+ extends this foundation with several key enhancements aimed at cultivar biology.
+These include a genome browser, gene model visualizer, feature search tools, and barcode generation utilities—together offering a more powerful and versatile platform for crop genetics research.
+
+%This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
 It outlines the integration of multiple extension features, providing the user with additional display and data analysis capabilities.
@@
 \section{Architecture}
-The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
+The \textit{Tersect Browser+} application is composed of four main components: 
+\begin{itemize}
+    \item the frontend; 
+    \item the backend; 
+    \item common utilities; and, 
+    \item extension modules.
+\end{itemize}
+
+The frontend is a browser-based interface developed in Angular and React, designed for interactive data visualization, including distance plots, accession selectors, and genome browser integration.
+The backend, built in Node.js and Express, provides API endpoints, data models, and utility scripts to serve genomic data and computation results from a MongoDB database.
+A set of common modules supports both frontend and backend logic with shared resources.
+Additionally, the architecture supports pluggable extensions, such as the genome-browser, which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons.
+Supporting data files, environment configurations, and Python scripting for preprocessing or bioinformatics workflows are organized in separate folders for data loading and virtual environments, enabling efficient pipeline integration and deployment.
+
+%The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
 This `bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
@@
 
-Extensions added to TersectBrowser+ include the integration of a genome browser capable of displaying a variety of different tracks, notably variants and gene models, search functionality to identify and display regions containing high impact variants, and a tool to generate accession-specific barcodes.
+%Extensions added to TersectBrowser+ include the integration of a genome browser capable of displaying a variety of different tracks, notably variants and gene models, search functionality to identify and display regions containing high impact variants, and a tool to generate accession-specific barcodes.
 These extensions are hosted in a separate section of the project, but interact with both frontent and backend services to provide a seamless user experience.
+
+\textit{Tersect Browser+} is primarily built using Angular 8, preserving the architecture of the original application.
+To enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22.
+This modular, "bolt-on" extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend and backend technologies.
+Extensions are developed independently and published to NPM, making them easily linkable through project dependencies.
+This hybrid approach ensures forward compatibility and scalability, while maintaining stability in the core application.
+Future versions of \textit{Tersect Browser+} may explore fully migrating the main application to a more unified and modern framework.
 

### 2025-05-05 TechDoc/TechDoc.tex
Includes some additional changes to the Introduction/Overview section

@@
 
-\author{David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}
+% Have put them in alphabetical order
+\author{Gabrielle Baumberg, Gregory Lupton, David Oluwasusi, Tanya Stead \\ Supervised by: Dr. Tomasz Kurowski}
 
@@
 
+The requirements for the application extension were described in [reference Tom’s briefing on this from March]. For the biologist to be able to:
+
+\begin{itemize}
+    \item open a genome browser within Tersect Browser and investigate specific variants.
+    \item see specific genes of interests against suspected introgressions. 
+    \item identify areas where high-impact variants are visible. 
+    \item be able to genereate DNA barcodes for the varioius cultivars and species in the dataset. 
+    \item to be able to load additional datasets, including for other crops. 
+\end{itemize}
+
+The extensions were designed and added by the named authors under the supervision of of Dr Tomasz Kurowski between March and May 2025.
+
 %This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
@@
 \end{itemize}
+
+[Insert figure in here]
 
@@
 \section{Deployment}
-The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment.
-Here, we expand upon these instructions and give a detailed overview of the structure of each extension.
+The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment. This section provides additional information.
+
+%Here, we expand upon these instructions and give a detailed overview of the structure of each extension.
 

### 2025-05-05 TechDoc/TechDoc.tex
Edits to tech doc

@@
     \item open a genome browser within Tersect Browser and investigate specific variants.
-    \item see specific genes of interests against suspected introgressions. 
+    \item visualise gene models in areas of suspected introgressions.
     \item identify areas where high-impact variants are visible. 
@@
 TersectBrowser+ has been tested with a dataset size of ~500 VCF files representing individual accessions.
+VCF files should be produced using SNPeff software.
+If not using SNPeff, ensure that variant impact levels are categorised into `High', `Moderate', and `Low' impact.
+If not provided, the gene model track in the Variant Browser will be unavailable.
     \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
-    \item A GFF file of gene model information for the reference genome, produced using SNPeff software.
-If not using SNPeff, ensure that variant impact levels are categorised into `High', `Medium', and `Low' impact.
-If not provided, the gene model track in the Variant Browser will be unavailable.
+    \item A GFF file of gene model information for the reference genome. 
     \item Whether there are prior identified introgressions known to be functionally relevant for the species.
@@
 Finally, callHighlightBins() is called, which itself calls highlightBins() from the bin-draw.service, passing as arguments the current interval start position, binsize, displayed accessions in an ordered format, and the searched accessions.
+
 \paragraph{\hl{Clear Button }}

### 2025-05-05 TechDoc/TechDoc.tex
Added Use Case section, Changed Section 4 title (now 5) to Features and Capabilities rather than Tersect Browser+ and other small polishing

--- 2025-05-05 TechDoc/TechDoc.tex
@@
 
-The requirements for the application extension were described in [reference Tom’s briefing on this from March]. For the biologist to be able to:
-
-\begin{itemize}
-    \item open a genome browser within Tersect Browser and investigate specific variants.
-    \item see specific genes of interests against suspected introgressions. 
-    \item identify areas where high-impact variants are visible. 
-    \item be able to genereate DNA barcodes for the varioius cultivars and species in the dataset. 
-    \item to be able to load additional datasets, including for other crops. 
-\end{itemize}
+The requirements for the application extension were described in [reference Tom’s briefing on this from March].
+For the biologist to be able to be able to explore plant genomes and introgressions in real time, by applying a set of flexible expressions, resequenced variant data and enabling dynamic visualisations.
 
@@
 
-[Insert figure in here]
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=15cm, height=9cm]{high_level_arch_draft.png}
+    \caption{Summary Architecture of TersectBrowser+.}
+\end{figure}
 
@@
 Future versions of \textit{Tersect Browser+} may explore fully migrating the main application to a more unified and modern framework.
+
+\section{Use Case}
+
+The study of cultivar genomes has evolved rapidly, with growing emphasis on mapping introgressions—segments of DNA from wild plant species that have been incorporated into cultivated lines.
+A variety of bioinformatic tools explore and characterize existing introgressions and identify commercially relevant regions for future breeding.
+Biologists have a requirement to explore plant genomes and introgressions in real time, by applying a set of flexible expressions, resequenced variant data and enabling dynamic visualisations.
+The requirements also include the ability to undertake such investigations with tools to browse genomes, view gene models, search for features and generate barcodes without leaving the web-page.
+
+The primary use cases considered were therefore the ability to:
+
+\begin{itemize}
+    \item (1) open a genome browser within Tersect Browser and investigate specific variants;
+    \item (2) see specific genes of interests against suspected introgressions; 
+    \item (3) identify areas where high-impact variants are visible; 
+    \item (4) be able to genereate DNA barcodes for the varioius cultivars and species in the dataset; and,  
+    \item (5) to be able to load additional datasets, including for other crops. 
+\end{itemize}
+
+%listed four main extension requirements, as well as a stretch extension.
+These requirements were provided as discrete goals for the project, and informed the main working priorities for the project timeline, however in practice the four extensions do have significant overlap in terms of code structure.
+For this reason, the accompanying TersectBrowser+ Article includes the extensions Genome Browser and Gene Models in the same section, while this Technical Documentation keeps the two separate according to the chronological timeline of developments to these extensions.
+Architecture diagrams are used to display the overlap between various extension components, especially relating to how they orchestrate data flow through the entire software.
 
@@
 
-The Admin is defined as an individual with bioinformatics and software design knowledge, background comprehension for all sections of this Technical Documentation, and acting to facilitate the use of a deployed TersectBrowser+ by the plant breeder main user of the software.
-When a new dataset is requested for deployment on TersectBrowser+, the Admin will need to follow the below steps:
+The Admin is defined as an individual with bioinformatics and software design knowledge, background comprehension for all sections of this Technical Documentation, and acting to facilitate the use of a deployed \textit{TersectBrowser+} by the plant breeder main user of the software.
+When a new dataset is requested for deployment on  \textit{TersectBrowser+}, the Admin will need to follow the below steps:
 \begin{enumerate}
@@
 \end{itemize}
-    \item Clone the GitHub repository of TersectBrowser+ to the local machine.
+    \item Clone the GitHub repository of  \textit{TersectBrowser+} to the local machine.
 Follow the README from \hyperlink {https://github.com/Tersect-Browser/Tersect-browser.git}{TersectBrowser+ GitHub} for installation instructions.
@@
 
-\section{TersectBrowser+}
-The project brief to extend the original TersectBrowser listed four main extension requirements, as well as a stretch extension.
-These requirements were provided as discrete goals for the project, and informed the main working priorities for the project timeline, however in practice the four extensions do have significant overlap in terms of code structure.
-For this reason, the accompanying TersectBrowser+ Article includes the extensions Genome Browser and Gene Models in the same section, while this Technical Documentation keeps the two separate according to the chronological timeline of developments to these extensions.
-Architecture diagrams are used to display the overlap between various extension components, especially relating to how they orchestrate data flow through the entire software.
+\section{Functional Capabilities and Extensions}
+
+\textit{Tersect Browser+} comprises Tersect CLI, Tersect Browser and four extensions that meet the \textit{Teresect Browser+} use case in Section 3. These were:
+
+\begin{itemize}
+    \item Genome Browser; 
+    \item Gene Model; 
+    \item Feature Search; and, 
+    \item Barcode Generation
+\end{itemize}
+
 \\
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=15cm, height=9cm]{high_level_arch_draft.png}
-    \caption{Summary Architecture of TersectBrowser+.}
-\end{figure}
 

### 2025-05-05 TechDoc/TechDoc.tex
Add to future improvements

--- 2025-05-05 TechDoc/TechDoc.tex
@@
 \subsubsection{Future Improvements}
+The gene models track could be expanded to display alternative splicing events, offering a clearer view of transcript diversity.
+Additionally, adding a hover-based popup over each bin that displays gene models for accessions containing variants could offer a rapid and informative view of variant differences.
 

### 2025-05-05 TechDoc/TechDoc.tex
Polishing of the Gene Browser and Gene Model sections

--- 2025-05-05 TechDoc/TechDoc.tex
@@
     \item Feature Search; and, 
-    \item Barcode Generation
+    \item Barcode Generation.
 \end{itemize}
@@
 The main benefit of TersectBrowser relative to similar genome comparison software for displaying key differences within resequenced datasets is its use of Tersect CLI, to index and then quickly query these genomes for investigation at user-defined scale.
-However, in order for TersectBrowser+ to be competitive as a tool with other browsers, it was necessary to implement features that enhance further investigation within the scope of the same browser.
-The main view of the original TersectBrowser displayed the density of variants along each chromosome in comparison to the reference genome.
-To enhance this view, a variant browser with multiple tracks providing extra information at a per-base level was added to enable to user to directly correlate variant density with accession-specific variants.
+To meet the user requirements and be competitive as a tool with other browsers, it was necessary to implement features that enable the visualised variant results to be annotated with genome information in the same browser.
+The main view of the original browser displayed the density of variants along each chromosome in comparison to the reference genome.
+To enhance this view, a variant browser with multiple tracks providing extra information at a per-base level was added.
+This enabled the user to directly correlate variant density with accession-specific variants.
 
-After a review of node-compatible browser tools such as CRAMER, JBrowse was ultimately chosen to carry out this extension request, due to its high customisability, up-to-date features, and active technical support.
-The initial build of JBrowse into TersectBrowser+ as an extension focused on retaining features most relevant to the user aims, and bringing the styling of the panel in line with the rest of the page.
-This extension also involved generating a popup window of the same JBrowse component to be initialised when the user was interested in a specific accession or bin region.
+On reviewing node-compatible browser tools such as CRAMER, JBrowse was selected as the prefered solution, due to its high customisability, up-to-date features and active technical support.
+The initial build of JBrowse into \textit{TersectBrowser+}, focused on retaining features most relevant to the user aims.
+This includeed bringing the styling of the panel in line with the rest of the browser page.
+The extension also involved generating a popup window of the same JBrowse component to be initialised when the user identified a specific accession or bin region of interest.
 
@@
 
-\paragraph{Feature Description}
-\begin{itemize}
-    \item JBrowse window:
-    \begin{itemize}
-        \item Main window is located at the top of Tersect Browser, below settings bar but above the heatmap 
-        \item Main window shows only sequence tracks and linear genome scale.
+\subsubsection{Main Window Location and Tracks}
+The main window is located at the top of Tersect Browser, below the settings bar but above the heatmap.
+Main window shows only sequence tracks and linear genome scale.
 Interactive features such as zoom anc chromosome selection have been removed from the main window.
 The container window has a dynamic offset, which changes according to tree view and zoom level to ensure the chromosome scales between TersectBrowser and JBrowse line up.
-        \item By default, the main window shows only the gene models track.
-        \item Popup window provides a more sandbox experience, providing more user interaction.
-The user can change zoom level and displayed chromosome, and two scale bars enable both precise and rapid navigation.
-The hamburger menu on the top left provides the capability for the user to upload their own data to visualise unique tracks.
-        \item By default, the popup window shows the reference sequence track, gene models track, and variant track for the specified accession.
-    \item Tracks:
+By default, the main window shows only the gene models track.
+
+The tracks include: 
 \begin{itemize}
@@
 \end{itemize}
-    \item Syncing of the JBrowse window and the TersectBrowser heatmap includes the following:
+
+\subsubsection{Popups and User Interaction}
+Popup window provides a more sandbox experience, providing more user interaction.
+The user can change zoom level and displayed chromosome, and two scale bars enable both precise and rapid navigation.
+The hamburger menu on the top left provides the capability for the user to upload their own data to visualise unique tracks.
+By default, the popup window shows the reference sequence track, gene models track, and variant track for the specified accession.
+
+\subsubsection{Jbrowser TersectBrowser Syncing}
+
+Syncing of the JBrowse window and the TersectBrowser heatmap includes the following:
     \begin{itemize}
@@
     \end{itemize}
-\end{itemize}
-\end{itemize}
 
-\subsubsection{Technical Description}
+\subsubsection{Technical Components}
 
@@
 The \verb +JBrowseLinearGenomeView+ component is imported from @jbrowse/react-linear-genome-view and is wrapped using React.
-The configurations for the assembly, assembly, and variant tracks, tracks, are imported from the respective assembly.ts and tracks.ts configuration files.
+The configurations for the assembly, assembly, and variant tracks, and tracks, are imported from the respective assembly.ts and tracks.ts configuration files.
 The styling configurations, config, are imported from jbrowseConfig.ts.
@@
 
-The method \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+.
+The method \verb +JbrowserWrapper+ takes as the arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+.
 If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the \verb +JbrowseWithAccessionName+ is returned (See below parameters).
@@
 If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+.
-This takes as arguments the following variables:
+This takes the following variables as arguments:
 \begin{itemize}
@@
 \paragraph{Horizontal Scroll}
-Horizontal scroll inputs from the user are disabled inside the main JBrowse window through the implementation of a \verb+Plugin+ of the \verb+DisableScroll+ class in \verb+JbrowserWrapper.tsx+, which overrides the JBrowse configuration to prevent horizontal scrolling
+Horizontal scroll inputs from the user are disabled inside the main JBrowse window through the implementation of a \verb+Plugin+ of the \verb+DisableScroll+ class in \verb+JbrowserWrapper.tsx+, which overrides the JBrowse configuration to prevent horizontal scrolling.
 
 \subsubsection{Test Results}
-When additional variant tracks are added to the Browser view, the size of the browser panel does not change. The user is able to scroll downwards to view more tracks in the browser.
+When additional variant tracks are added to the Browser view, the size of the browser panel does not change. The user is, however, able to scroll downwards to view more tracks in the browser.
 
-When a new dataset is added, the custom addition script automatically updates the names of tracks and assembly in the relevant files within the genome-browser extension.
+When a new dataset is added, the custom additional script automatically updates the names of tracks and assembly in the relevant files within the genome-browser extension.
 This flexibility allows the Genome Browser extension to be accessible by plant breeders working on many different datasets.
 TersectBrowser was initially validated with a human genome dataset, so it was important to maintain this flexibility of input in the new extensions.
-
-\subsubsection{Future Improvements}
-In addition to displaying the reference track, gene models track, and variants track, additional feature tracks, such as a multiple sequence alignment view, could be added to the genome browser.
-
-Currently, the genome browser popup window is limited to opening a single window within the TersectBrowser+ webpage.
-A button to open the popup window in a new tab could be added, which would allow the user to open multiple detatched popup windows and view them simultaneously.
 
@@
 \subsubsection{Rationale}
-Once the Genome Browser panel was added to the main page of TersectBrowser+, this facilitated the inclusion of a reference gene model track.
+Once the Genome Browser panel was added to the main page of TersectBrowser+, this facilitated the inclusion of a reference gene model track using the same techical solution.
 This allows the plant breeder user to identify an area of high variant density on the heatmap, and then check the gene model track to investigate whether this region occurs near a gene of interest.
@@
 \subsubsection{User Interface}
-\paragraph{Feature Description}
-\begin{itemize}
-    \item The gene models track is displayed in the main JBrowse window. 
-    \item In the popup window, the gene models track is displayed as the second track, underneath the reference sequence track and above the variant track.
-\end{itemize}
 
-\subsubsection{Technical Description}
+[Insert a figure]
+
+\subsubsection{Alignment with JBrowser}
+The gene models track is displayed in the main JBrowse window.
+In the popup window, the gene models track is displayed as the second track, underneath the reference sequence track and above the variant track.
+
+\subsubsection{Technical Components}
 The configuration for the Gene Models \verb+FeatureTrack+ is defined as the first entry in \verb+tracks.ts+.
@@
 The relevance of the gene models in the GFF file depend on what the user provides in the dataset, and how the GFF file was created.
-
-\subsubsection{Future Improvements}
 
@@
 \subsubsection{Test Results}
-
-\subsubsection{Future Improvements}
-Currently, only bins on the TersectBrowser+ heatmap are highlighted in red to indicate the positions of high-impact variants.
-Highlighting could be extended across the Genome Browser and Gene Models extension, so that gene models matching the parameters of the feature search query are also highlighted in red.
 
@@
 \subsubsection{Test Results}
-\subsubsection{Future Improvements}
+
+\subsection{Future Improvements}
+
+\subsubsection{Gene Browser}
+In addition to displaying the reference track, gene models track, and variants track, additional feature tracks, such as a multiple sequence alignment view, could be added to the genome browser.
+
+Currently, the genome browser popup window is limited to opening a single window within the TersectBrowser+ webpage.
+A button to open the popup window in a new tab could be added, which would allow the user to open multiple detatched popup windows and view them simultaneously.
+
+\subsubsection{Feature Search}
+Currently, only bins on the TersectBrowser+ heatmap are highlighted in red to indicate the positions of high-impact variants.
+Highlighting could be extended across the Genome Browser and Gene Models extension, so that gene models matching the parameters of the feature search query are also highlighted in red.
+
+\subsubsection{Barcoding}
 Currently, barcode viewing is limited to the downloaded file. It can be extended to display barcodes in the popup window, with the variant highlighted in red.
 
-\subsection{Future Work and Extension Design}
 \subsubsection{Automated Introgression Search Extension}
+
+\section{Conclusions}
 

### 2025-05-05 TechDoc/TechDoc.tex

@@
 
+\subsection{Future Improvements}
+Currently, the genome browser popup window is limited to opening a single window within the TersectBrowser+ webpage.
+A button to open the popup window in a new tab could be added, which would allow the user to open multiple detatched popup windows and view them simultaneously.
+
 \subsection{Gene Models Extension}
@@
 The relevance of the gene models in the GFF file depend on what the user provides in the dataset, and how the GFF file was created.
+
+\subsubsection{Future Improvements}
+The gene models track could be expanded to display alternative splicing events, offering a clearer view of transcript diversity.
+Additionally, adding a hover-based popup over each bin that displays gene models for accessions containing variants could offer a rapid and informative view of variant differences.
 

### 2025-05-05 TechDoc/TechDoc.tex
upload feature search text

--- 2025-05-05 TechDoc/TechDoc.tex
@@
 
-% Have put them in alphabetical order
-\author{Gabrielle Baumberg, Gregory Lupton, David Oluwasusi, Tanya Stead \\ Supervised by: Dr. Tomasz Kurowski}
+\author{David Oluwasusi, Tanya Stead, Gregory Lupton, Gabrielle Baumberg \\ Supervised by: Dr. Tomasz Kurowski}
 
@@
 \section{Overview}
-This document describes the technical specifications and extended functionalities of Tersect Browser+, a web-based application for the exploration of genome and variant data in cultivated and wild-type crops.
-Building upon the original Tersect Browser, developed by Tomasz Kurowski [Reference], this enhanced version integrates additional features to support the identification and analysis of introgressions within accession species.
-The system leverages a MongoDB database to manage genomic datasets efficiently and provides a suite of interactive tools for biological interpretation.
-
-The original Tersect Browser offered a streamlined interface for investigating genomic intervals and phylogenetic relationships, powered by the high-performance Tersect CLI.
-It included dynamic visualizations such as phylogenetic trees and genetic distance heatmaps derived from resequenced genome data.
-
-Tersect Browser+ extends this foundation with several key enhancements aimed at cultivar biology.
-These include a genome browser, gene model visualizer, feature search tools, and barcode generation utilities—together offering a more powerful and versatile platform for crop genetics research.
-
-The requirements for the application extension were described in [reference Tom’s briefing on this from March]. For the biologist to be able to:
-
-\begin{itemize}
-    \item open a genome browser within Tersect Browser and investigate specific variants.
-    \item visualise gene models in areas of suspected introgressions.
-    \item identify areas where high-impact variants are visible. 
-    \item be able to genereate DNA barcodes for the varioius cultivars and species in the dataset. 
-    \item to be able to load additional datasets, including for other crops. 
-\end{itemize}
-
-The extensions were designed and added by the named authors under the supervision of of Dr Tomasz Kurowski between March and May 2025.
-
-%This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
+This Technical Documentation describes the particular details of building upon the original TersectBrowser developed by Tomasz Kurowski into  TersectBrowser+.
 It outlines the integration of multiple extension features, providing the user with additional display and data analysis capabilities.
@@
 \section{Architecture}
-The \textit{Tersect Browser+} application is composed of four main components: 
-\begin{itemize}
-    \item the frontend; 
-    \item the backend; 
-    \item common utilities; and, 
-    \item extension modules.
-\end{itemize}
-
-[Insert figure in here]
-
-The frontend is a browser-based interface developed in Angular and React, designed for interactive data visualization, including distance plots, accession selectors, and genome browser integration.
-The backend, built in Node.js and Express, provides API endpoints, data models, and utility scripts to serve genomic data and computation results from a MongoDB database.
-A set of common modules supports both frontend and backend logic with shared resources.
-Additionally, the architecture supports pluggable extensions, such as the genome-browser, which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons.
-Supporting data files, environment configurations, and Python scripting for preprocessing or bioinformatics workflows are organized in separate folders for data loading and virtual environments, enabling efficient pipeline integration and deployment.
-
-%The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
-This `bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
+The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22.
+This 'bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts.
 Additionally, this approach allows the new extnesions to use up-to-date packages, resulting in enhanced functionality and a modern design.
@@
 
-%Extensions added to TersectBrowser+ include the integration of a genome browser capable of displaying a variety of different tracks, notably variants and gene models, search functionality to identify and display regions containing high impact variants, and a tool to generate accession-specific barcodes.
+Extensions added to TersectBrowser+ include the integration of a genome browser capable of displaying a variety of different tracks, notably variants and gene models, search functionality to identify and display regions containing high impact variants, and a tool to generate accession-specific barcodes.
 These extensions are hosted in a separate section of the project, but interact with both frontent and backend services to provide a seamless user experience.
 
-\textit{Tersect Browser+} is primarily built using Angular 8, preserving the architecture of the original application.
-To enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22.
-This modular, "bolt-on" extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend and backend technologies.
-Extensions are developed independently and published to NPM, making them easily linkable through project dependencies.
-This hybrid approach ensures forward compatibility and scalability, while maintaining stability in the core application.
-Future versions of \textit{Tersect Browser+} may explore fully migrating the main application to a more unified and modern framework.
-
 \section{Deployment}
-The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment. This section provides additional information.
-
-%Here, we expand upon these instructions and give a detailed overview of the structure of each extension.
+The GitHub project \url{https://github.com/Tersect-Browser/Tersect-browser} README provides detailed instructions for setup, dependencies, and deployment.
+Here, we expand upon these instructions and give a detailed overview of the structure of each extension.
 
@@
 TersectBrowser+ has been tested with a dataset size of ~500 VCF files representing individual accessions.
-VCF files should be produced using SNPeff software.
-If not using SNPeff, ensure that variant impact levels are categorised into `High', `Moderate', and `Low' impact.
+    \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
+    \item A GFF file of gene model information for the reference genome, produced using SNPeff software.
+If not using SNPeff, ensure that variant impact levels are categorised into 'High', 'Medium', and 'Low' impact.
 If not provided, the gene model track in the Variant Browser will be unavailable.
-    \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
-    \item A GFF file of gene model information for the reference genome. 
     \item Whether there are prior identified introgressions known to be functionally relevant for the species.
-These can be used to `check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
+These can be used to 'check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
 \end{itemize}
@@
 \subsubsection{User Interface}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=10cm, height=8cm]{GenomeBrowser.png}
-    \caption{UI diagram for Genome Browser.}
-\end{figure}
 
@@
 \end{itemize}
-\end{itemize}
 
@@
 If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated.
-If either of them are empty, a container with the phrase ``Loading…" is returned.
+If either of them are empty, a container with the phrase "Loading…" is returned.
 If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+.
@@
 The configurations for the assembly track, which is built upon the reference accession SL2.50, is defined in \verb +assembly.ts+ in json format.
-The name is set to the reference accession name and \verb +trackId+ is set to ``SL2.50-ReferenceSequenceTrack".
+The name is set to the reference accession name and \verb +trackId+ is set to "SL2.50-ReferenceSequenceTrack".
 The paths to the fasta file, along with its corresponding fasta index, are specified as local server URLs provided by the backend during the Tersect Browser setup.
@@
 The configurations for the \verb +VariantTrack+ and Gene Models \verb +FeatureTrack+ are defined in \verb +tracks.ts+ in json format.
-For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to ``ITAG2.4 Gene Models".
+For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to "ITAG2.4 Gene Models".
 The paths to the zipped files, along with its corresponding Tabix index files, are specified as local server URLs provided by the backend during the Tersect Browser setup.
@@
 \paragraph{Styling}
-The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour `\#459e00' and the \verb +boxShadow+ is set to `none'.
+The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour '\#459e00' and the \verb +boxShadow+ is set to 'none'.
 
@@
 \subsubsection{Future Improvements}
-The gene models track could be expanded to display alternative splicing events, offering a clearer view of transcript diversity.
-Additionally, adding a hover-based popup over each bin that displays gene models for accessions containing variants could offer a rapid and informative view of variant differences.
 
@@
 \subsubsection{User Interface}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=15cm, height=9cm]{FS1.png}
-    \caption{UI digram for Feature Search.}
-\end{figure}
-
 \paragraph{Feature Description}
@@
 \end{itemize}
-\subsubsection{Technical Description}
+\subsection*{4.3.3 Feature Search Extension – Technical Description}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=7cm, height=15cm]{FS2.png}
-    \caption{Data flow diagram for Feature Search.}
-\end{figure}
+\subsubsection*{4.3.3.1 Creating Search Indexes Per Chromosome}
 
-\paragraph{\hl{Search Bar}}
-\paragraph{\hl{Popup window with advanced settings }}
-\paragraph{Highlighting bins}
-Bin highlighting is controlled in the \verb+bin-draw.service+ by two functions.
-The first function, \verb+highlightFeatureBins()+, is defined, taking as arguments a string containing accession names, the bin position along the x-axis, and the \verb+binView+.
-First, the y-axis bin position for accession names is calculated.
-Then, the \verb+binView+ is redrawn in greyscale, with the bin colour determined by the difference to the reference accession.
-Using the x-axis \verb+binIndex+ position and the y-axis accession bin positions, these bins are coloured red in the \verb+binView+.
-The modified \verb+binView+ is returned as the output.
-The second function, \verb+highlightBins()+, takes as arguments the start position of the interval and the bin size currently shown in the Tersect pane, a list of ordered accessions shown in the canvas, and the searched accessions.
-For each searched accession, accession name is reformatted to match the format displayed in the Tersect pane.
-Then, the bin position along the x-axis is calculated using the \verb+getBinIndexFromPosition()+ function, which takes as arguments the feature position along the chromosome, the interval start position, and the bin size.
-Then, \verb+highlightFeatureBins()+ is called for each accession.
-Lastly, the canvas is redrawn using the modified \verb+binView+ output from \verb+highlightFeatureBins()+.
+To enable searching for high impact variants across the dataset, distinct variant sites per chromosome region were obtained using \texttt{tersect}.
+The resulting VCF file was annotated with the merged VCF files from all accession samples in the dataset.
+
+The final VCF was then compressed using \texttt{bgzip} and indexed with \texttt{tabix}.
+Additionally, Trix index files \citep{ucsc_trix} were generated for high impact variants on each chromosome to support rapid gene name search suggestions.
+
+\subsubsection*{4.3.3.2 Handling Search Request}
+
+An API endpoint was added to the \texttt{tersect-router} backend to support search functionality. The structure of the endpoint is as follows:
+
+\begin{verbatim}
+/query/:datasetId/variants/:chrom?start=number&end=number&filter=HIGH|MODERATE|LOW
+\end{verbatim}
+
+When searching by gene name, the Trix index files containing gene names associated with high impact variants are utilized.
+
+\subsubsection*{4.3.3.3 Handling Searching by Interval}
+
+The chromosome name is passed as a parameter to the search endpoint, while the start and end positions and the impact filter are passed as query parameters.
+
+The search functionality is powered by \texttt{bcftools}, which is executed using the \texttt{child\_process} module in \texttt{Node.js}.
+The matching variants are captured by parsing their identifiers in the format \texttt{CHROM:POS:REF:ALT}, and these are stored in a set to ensure uniqueness.
+
+To associate variants with accession names, a function uses \texttt{tersect} to determine which accessions have each specific variant.
+A function to get the optimal concurrency (\texttt(getOptimalConcurrency)) was introduced which detected the number of cores present on the server, subtracting 1 from the total and using the rest.
+To prevent server overload from concurrent \texttt{tersect} calls, the \texttt{p-limit} library is used to control concurrency based on available CPU cores.
+
+The positions are then collated to a map that links positions to a list of accession names.
+The results are then constructed by looping through each line and getting specific accession list from the map created via the position.
+
+The final result maps positions to accession name lists. These results are formatted and returned to the frontend in the following structure:
+
+\begin{verbatim}
+{
+  region,
+  filter,
+  count,
+  variants,
+  totalAccessions
+}
+\end{verbatim}
+
+The values are described below;
+
 \begin{itemize}
-    \item highlightFeatureBins(accessions: string[], binIndex: number, binView: DistanceBinView) - takes a string of accession names, a binIndex (corresponding to the bin position along the x-axis matching the gene position on the chromosome), and binView.
-The y-axis index for bins matching accession names in the accessions strings is calculated and combined with the binIndex to colour these specific bins red.
-The rest of the bins are coloured in greyscale, with saturation depending on binDistance (calculated from tersect on the backend
-    \begin{itemize}
-        \item Bin-draw.service.ts 
-        \item Called by highlightBins() in bin-draw.service.ts 
-    \end{itemize}
-    \item highlightBins(intervalStart, binsize, orderedAccessions, searchedAccessions) - takes selected interval start position, selected binSize, list of ordered accessions shown in the canvas, and searched accessions (passed from callback function?).
-Accession names in Jbrowse are in a different format to what is stored in tersect browser, so accession names are reformatted to match tersect browser bins.
-binIndex is calculated for the selected gene.
-These two are passed to highlightFeatureBins() - along with this.bins - to highlight the bins.
-The canvas is then redrawn using the binView calculated in highlightFeatureBins().
-    \begin{itemize}
-        \item Bin-draw.service.ts 
-        \item Called by callHighlightBins() in tersect-browser.component.ts 
-    \end{itemize}
-    \item callHighlightBins(searchedAcessions) - takes searchedAccessions.
-Calls highlightBins(), passing along  selected interval start position, selected binSize, list of ordered accessions shown in the canvas, and searched accessions (passed from callback function?).
-    \begin{itemize}
-        \item Tersect-browser.component.ts 
-    \end{itemize}
+    \item \textbf{region} - a string containing the region or interval on the selected chromosome (e.g., "chr1:100000-200000")
+    \item \textbf{filter} - a string specifying the variant impact filter applied (e.g., "HIGH", "MODERATE", or "LOW")
+    \item \textbf{count} - an integer representing the total number of variants found in the region with the specified filter
+    \item \textbf{variants} - an array of variant objects, each including chromosome, position, reference, alternate, effect, and accessions
+    \item \textbf{totalAccessions} - an integer specifying the total number of accessions in the dataset
 \end{itemize}
 
-\paragraph{Calling Highlighting Bins} \mbox{}
-\\
-\hl{TBC: Mechanism} of searching VCFs to identify variants
+Each variant is formatted as:
 
-Finally, callHighlightBins() is called, which itself calls highlightBins() from the bin-draw.service, passing as arguments the current interval start position, binsize, displayed accessions in an ordered format, and the searched accessions.
+\begin{verbatim}
+[
+  {
+    chr,
+    pos: { start: pos },
+    ref,
+    alt,
+    eff,
+    accessions
+  }
+]
+\end{verbatim}
 
-\paragraph{\hl{Clear Button }}
-The clear button is located to the right of the \verb+Open variant search+ button and styled with \verb+fa fa-refresh+.
-It contains the event handler \verb+(click)="refreshBin()"+ that calls the \verb+refreshBin()+ method in \verb+BinDrawService+.
-This method calls \verb+generatePlotArray()+, which re-creates the bin view, and \verb+updateCanvas()+, which redraws the canvas using the re-created bin view.
+The values are described below;
 
-\subsubsection{Test Results}
+\begin{itemize}
+    \item \textbf{chr} - a string specifying the chromosome name
+    \item \textbf{pos} - an object with a \texttt{start} field indicating the variant position
+    \item \textbf{ref} - a string specifying the reference allele
+    \item \textbf{alt} - a string specifying the alternate allele
+    \item \textbf{eff} - a string describing the effect or impact of the variant (e.g., "HIGH", "MODERATE", "LOW")
+    \item \textbf{accessions} - an array of accession names (strings) that contain this variant
+\end{itemize}
 
-\subsubsection{Future Improvements}
-Currently, only bins on the TersectBrowser+ heatmap are highlighted in red to indicate the positions of high-impact variants.
-Highlighting could be extended across the Genome Browser and Gene Models extension, so that gene models matching the parameters of the feature search query are also highlighted in red.
+\subsubsection*{4.3.3.4 Handling Searching by Gene Name}
+
+Gene-based searches use the pre-built Trix index files, which currently include only high impact variants.
+
+The Trix index is generated by exporting high impact variants per chromosome to a text file, using this identifier format:
+
+\begin{verbatim}
+docId = "${chrom}:${pos}:${ref}:${alt}|${sample}|${effBlock}"
+\end{verbatim}
+
+The index is built using \texttt{ixIxx}.
+
+\subsubsection*{4.3.3.5 Linking with the Frontend}
+
+The gene search extension is developed in React to enable integration with future \texttt{JBrowse} sessions.
+Despite the main application being Angular (v19), React components are embedded using the \texttt{ReactComponentDirective} from the \texttt{@ngeenx/ngx-react} package.
+
+Two UI components support search configuration:
+\begin{itemize}
+  \item A range slider to select intervals.
+  \item An autocomplete input for gene ID search.
+\end{itemize}
+
+The React component handles the form submission, triggers the appropriate search via API or Trix, and returns results through a callback.
+Bin highlighting is then managed by the legacy tersect browser app.
+
+\subsubsection*{4.3.3.6 Other Ways to Trigger Variant Search}
+
+Variant searches can also be triggered from:
+\begin{itemize}
+  \item Each bin on the bin plot.
+  \item A selected range on the scale plot.
+\end{itemize}
+
+These triggers emit a custom event \texttt{tersect-search-variants}, which sets the interval and opens the modal for user interaction.
+
+\subsubsection*{4.3.3.7 Highlighting Bins}
+
+Bin highlighting is handled by two main functions in \texttt{bin-draw.service.ts}:
+
+\begin{itemize}
+  \item \texttt{highlightFeatureBins(accessions, binIndex, binView)}: Computes the y-axis position for each accession, redraws the bin view in greyscale, and highlights selected bins in red.
+  \item \texttt{highlightBins(intervalStart, binSize, orderedAccessions, searchedAccessions)}: Calculates the \texttt{binIndex} from position and bin size, and calls \texttt{highlightFeatureBins()}.
+\end{itemize}
+
+The top-level function \texttt{callHighlightBins(searchedAccessions)} is defined in \texttt{tersect-browser.component.ts}, where it retrieves interval, bin size, and accession info, and calls \texttt{highlightBins()}.
+
+\subsubsection*{VCF Mechanism Summary}
+
+The overall mechanism for variant search begins with API queries, uses either \texttt{bcftools} or Trix for data retrieval, links variants to accessions with \texttt{tersect}, and finally triggers the highlighting of bins in the Tersect UI.
+
+\paragraph{Other Triggers for Variant Search}
+Variant search also triggered via bin plot or scale plot selection using the \texttt{tersect-search-variants} custom event.
+
+\clearpage
 
@@
     \centering
-    \includegraphics[width=6cm, height=8cm]{Bar1.png}
-    \caption{UI diagram for Barcode Generation.}
+    \includegraphics[width=5cm, height=9cm]{UserBarcode.jpg}
+    \caption{User-focused data flow through Barcoding Extension.}
 \end{figure}
-
 \paragraph{Frontend Display Feature Description}
-When the user clicks on a bin, the popup menu appears and the button ``Create barcode" is visible, styled with a barcode icon.
+When the user clicks on a bin, the popup menu appears and the button "Create barcode" is visible, styled with a barcode icon.
 When the user clicks on this button, a popup window appears in the centre of the screen.
 The window title is the corresponding accession name.
-The window contains the header ``Barcode Generator" and contains a description of what is being generated.
+The window contains the header "Barcode Generator" and contains a description of what is being generated.
 There is a box for the user to input barcode size, with default set to 150, and the maximum number of variants allowed per barcode (this is an optional parameter.
@@
 
-At the bottom is a green ``Generate" button.
+At the bottom is a green "Generate" button.
 Hovering over the button turns it dark green, and clicking on this button will call the backend scripts to generate the barcodes output.
 Whilst scripts are running and barcodes are being generated, a spinner is added to the button to indicate loading.
-After the script has been generated, the spinner disappears and another green button ``Download" appears at the bottom left of the window.
+After the script has been generated, the spinner disappears and another green button "Download" appears at the bottom left of the window.
 Clicking on the button downloads the tsv file containing the barcodes.
@@
     \centering
-    \includegraphics[width=7cm, height=15cm]{Bar2.png}
-    \caption{Data flow diagram for Barcode Generation.}
+    \includegraphics[width=7cm, height=12cm]{ExtBarcode.jpg}
+    \caption{User-focused data flow through Barcoding Extension.}
 \end{figure}
@@
 
-The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to ``True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+.
+The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to "True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+.
 The returned URL and filename are saved to the variables \verb+downloadUrl+ and \verb+downloadFileName+.
-Once returned, \verb+isLoading+ is set to ``False" and \verb+barcodeSize+ is set back to ``150".
+Once returned, \verb+isLoading+ is set to "False" and \verb+barcodeSize+ is set back to "150".
 
-The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to ``150``, \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the ``Download" button from the popup.
+The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to "150", \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the "Download" button from the popup.
 
@@
 The popup window is visible depending on the modal state isVisible, and is centrally position with \verb+modalTitle+ set as the header.
-Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button ``Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated.
-An anchor displays a clickable ``Download" p-button that points towards the generated barcode file.
-The filename is set to \verb+downloadFileName+, and ``Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.
+Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button "Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated.
+An anchor displays a clickable "Download" p-button that points towards the generated barcode file.
+The filename is set to \verb+downloadFileName+, and "Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.
 
@@
 
-The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+`TB_Barcode_Gen'+, and the specified accession.
+The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+'TB_Barcode_Gen'+, and the specified accession.
 
@@
 \section {References}
-
+\begin{thebibliography}{9}
+\bibitem{ucsc_trix}
+University of California, Santa Cruz. (n.d.). Trix Index Files. UCSC Genome Browser.
+\end{thebibliography}
 \end{document}

### 2025-05-05 TechDoc/TechDoc.tex
merge main

--- 2025-05-05 TechDoc/TechDoc.tex
@@
 \end{itemize}
-\subsubsection{Technical Description}
+\subsection*{4.3.3 Feature Search Extension – Technical Description}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=7cm, height=15cm]{FS2.png}
-    \caption{Data flow diagram for Feature Search.}
-\end{figure}
+\subsubsection*{4.3.3.1 Creating Search Indexes Per Chromosome}
 
-\paragraph{\hl{Search Bar}}
-\paragraph{\hl{Popup window with advanced settings }}
-\paragraph{Highlighting bins}
-Bin highlighting is controlled in the \verb+bin-draw.service+ by two functions.
-The first function, \verb+highlightFeatureBins()+, is defined, taking as arguments a string containing accession names, the bin position along the x-axis, and the \verb+binView+.
-First, the y-axis bin position for accession names is calculated.
-Then, the \verb+binView+ is redrawn in greyscale, with the bin colour determined by the difference to the reference accession.
-Using the x-axis \verb+binIndex+ position and the y-axis accession bin positions, these bins are coloured red in the \verb+binView+.
-The modified \verb+binView+ is returned as the output.
-The second function, \verb+highlightBins()+, takes as arguments the start position of the interval and the bin size currently shown in the Tersect pane, a list of ordered accessions shown in the canvas, and the searched accessions.
-For each searched accession, accession name is reformatted to match the format displayed in the Tersect pane.
-Then, the bin position along the x-axis is calculated using the \verb+getBinIndexFromPosition()+ function, which takes as arguments the feature position along the chromosome, the interval start position, and the bin size.
-Then, \verb+highlightFeatureBins()+ is called for each accession.
-Lastly, the canvas is redrawn using the modified \verb+binView+ output from \verb+highlightFeatureBins()+.
+To enable searching for high impact variants across the dataset, distinct variant sites per chromosome region were obtained using \texttt{tersect}.
+The resulting VCF file was annotated with the merged VCF files from all accession samples in the dataset.
+
+The final VCF was then compressed using \texttt{bgzip} and indexed with \texttt{tabix}.
+Additionally, Trix index files \citep{ucsc_trix} were generated for high impact variants on each chromosome to support rapid gene name search suggestions.
+
+\subsubsection*{4.3.3.2 Handling Search Request}
+
+An API endpoint was added to the \texttt{tersect-router} backend to support search functionality. The structure of the endpoint is as follows:
+
+\begin{verbatim}
+/query/:datasetId/variants/:chrom?start=number&end=number&filter=HIGH|MODERATE|LOW
+\end{verbatim}
+
+When searching by gene name, the Trix index files containing gene names associated with high impact variants are utilized.
+
+\subsubsection*{4.3.3.3 Handling Searching by Interval}
+
+The chromosome name is passed as a parameter to the search endpoint, while the start and end positions and the impact filter are passed as query parameters.
+
+The search functionality is powered by \texttt{bcftools}, which is executed using the \texttt{child\_process} module in \texttt{Node.js}.
+The matching variants are captured by parsing their identifiers in the format \texttt{CHROM:POS:REF:ALT}, and these are stored in a set to ensure uniqueness.
+
+To associate variants with accession names, a function uses \texttt{tersect} to determine which accessions have each specific variant.
+A function to get the optimal concurrency (\texttt(getOptimalConcurrency)) was introduced which detected the number of cores present on the server, subtracting 1 from the total and using the rest.
+To prevent server overload from concurrent \texttt{tersect} calls, the \texttt{p-limit} library is used to control concurrency based on available CPU cores.
+
+The positions are then collated to a map that links positions to a list of accession names.
+The results are then constructed by looping through each line and getting specific accession list from the map created via the position.
+
+The final result maps positions to accession name lists. These results are formatted and returned to the frontend in the following structure:
+
+\begin{verbatim}
+{
+  region,
+  filter,
+  count,
+  variants,
+  totalAccessions
+}
+\end{verbatim}
+
+The values are described below;
+
 \begin{itemize}
-    \item highlightFeatureBins(accessions: string[], binIndex: number, binView: DistanceBinView) - takes a string of accession names, a binIndex (corresponding to the bin position along the x-axis matching the gene position on the chromosome), and binView.
-The y-axis index for bins matching accession names in the accessions strings is calculated and combined with the binIndex to colour these specific bins red.
-The rest of the bins are coloured in greyscale, with saturation depending on binDistance (calculated from tersect on the backend
-    \begin{itemize}
-        \item Bin-draw.service.ts 
-        \item Called by highlightBins() in bin-draw.service.ts 
-    \end{itemize}
-    \item highlightBins(intervalStart, binsize, orderedAccessions, searchedAccessions) - takes selected interval start position, selected binSize, list of ordered accessions shown in the canvas, and searched accessions (passed from callback function?).
-Accession names in Jbrowse are in a different format to what is stored in tersect browser, so accession names are reformatted to match tersect browser bins.
-binIndex is calculated for the selected gene.
-These two are passed to highlightFeatureBins() - along with this.bins - to highlight the bins.
-The canvas is then redrawn using the binView calculated in highlightFeatureBins().
-    \begin{itemize}
-        \item Bin-draw.service.ts 
-        \item Called by callHighlightBins() in tersect-browser.component.ts 
-    \end{itemize}
-    \item callHighlightBins(searchedAcessions) - takes searchedAccessions.
-Calls highlightBins(), passing along  selected interval start position, selected binSize, list of ordered accessions shown in the canvas, and searched accessions (passed from callback function?).
-    \begin{itemize}
-        \item Tersect-browser.component.ts 
-    \end{itemize}
+    \item \textbf{region} - a string containing the region or interval on the selected chromosome (e.g., "chr1:100000-200000")
+    \item \textbf{filter} - a string specifying the variant impact filter applied (e.g., "HIGH", "MODERATE", or "LOW")
+    \item \textbf{count} - an integer representing the total number of variants found in the region with the specified filter
+    \item \textbf{variants} - an array of variant objects, each including chromosome, position, reference, alternate, effect, and accessions
+    \item \textbf{totalAccessions} - an integer specifying the total number of accessions in the dataset
 \end{itemize}
 
-\paragraph{Calling Highlighting Bins} \mbox{}
-\\
-\hl{TBC: Mechanism} of searching VCFs to identify variants
+Each variant is formatted as:
 
-Finally, callHighlightBins() is called, which itself calls highlightBins() from the bin-draw.service, passing as arguments the current interval start position, binsize, displayed accessions in an ordered format, and the searched accessions.
+\begin{verbatim}
+[
+  {
+    chr,
+    pos: { start: pos },
+    ref,
+    alt,
+    eff,
+    accessions
+  }
+]
+\end{verbatim}
 
-\paragraph{\hl{Clear Button }}
-The clear button is located to the right of the \verb+Open variant search+ button and styled with \verb+fa fa-refresh+.
-It contains the event handler \verb+(click)="refreshBin()"+ that calls the \verb+refreshBin()+ method in \verb+BinDrawService+.
-This method calls \verb+generatePlotArray()+, which re-creates the bin view, and \verb+updateCanvas()+, which redraws the canvas using the re-created bin view.
+The values are described below;
 
-\subsubsection{Test Results}
+\begin{itemize}
+    \item \textbf{chr} - a string specifying the chromosome name
+    \item \textbf{pos} - an object with a \texttt{start} field indicating the variant position
+    \item \textbf{ref} - a string specifying the reference allele
+    \item \textbf{alt} - a string specifying the alternate allele
+    \item \textbf{eff} - a string describing the effect or impact of the variant (e.g., "HIGH", "MODERATE", "LOW")
+    \item \textbf{accessions} - an array of accession names (strings) that contain this variant
+\end{itemize}
+
+\subsubsection*{4.3.3.4 Handling Searching by Gene Name}
+
+Gene-based searches use the pre-built Trix index files, which currently include only high impact variants.
+
+The Trix index is generated by exporting high impact variants per chromosome to a text file, using this identifier format:
+
+\begin{verbatim}
+docId = "${chrom}:${pos}:${ref}:${alt}|${sample}|${effBlock}"
+\end{verbatim}
+
+The index is built using \texttt{ixIxx}.
+
+\subsubsection*{4.3.3.5 Linking with the Frontend}
+
+The gene search extension is developed in React to enable integration with future \texttt{JBrowse} sessions.
+Despite the main application being Angular (v19), React components are embedded using the \texttt{ReactComponentDirective} from the \texttt{@ngeenx/ngx-react} package.
+
+Two UI components support search configuration:
+\begin{itemize}
+  \item A range slider to select intervals.
+  \item An autocomplete input for gene ID search.
+\end{itemize}
+
+The React component handles the form submission, triggers the appropriate search via API or Trix, and returns results through a callback.
+Bin highlighting is then managed by the legacy tersect browser app.
+
+\subsubsection*{4.3.3.6 Other Ways to Trigger Variant Search}
+
+Variant searches can also be triggered from:
+\begin{itemize}
+  \item Each bin on the bin plot.
+  \item A selected range on the scale plot.
+\end{itemize}
+
+These triggers emit a custom event \texttt{tersect-search-variants}, which sets the interval and opens the modal for user interaction.
+
+\subsubsection*{4.3.3.7 Highlighting Bins}
+
+Bin highlighting is handled by two main functions in \texttt{bin-draw.service.ts}:
+
+\begin{itemize}
+  \item \texttt{highlightFeatureBins(accessions, binIndex, binView)}: Computes the y-axis position for each accession, redraws the bin view in greyscale, and highlights selected bins in red.
+  \item \texttt{highlightBins(intervalStart, binSize, orderedAccessions, searchedAccessions)}: Calculates the \texttt{binIndex} from position and bin size, and calls \texttt{highlightFeatureBins()}.
+\end{itemize}
+
+The top-level function \texttt{callHighlightBins(searchedAccessions)} is defined in \texttt{tersect-browser.component.ts}, where it retrieves interval, bin size, and accession info, and calls \texttt{highlightBins()}.
+
+\subsubsection*{VCF Mechanism Summary}
+
+The overall mechanism for variant search begins with API queries, uses either \texttt{bcftools} or Trix for data retrieval, links variants to accessions with \texttt{tersect}, and finally triggers the highlighting of bins in the Tersect UI.
+
+\paragraph{Other Triggers for Variant Search}
+Variant search also triggered via bin plot or scale plot selection using the \texttt{tersect-search-variants} custom event.
+
+\clearpage
 

### 2025-05-05 TechDoc/TechDoc.tex
add search extension docs

@@
 
-\subsubsection{Future Improvements}
-The gene models track could be expanded to display alternative splicing events, offering a clearer view of transcript diversity.
-Additionally, adding a hover-based popup over each bin that displays gene models for accessions containing variants could offer a rapid and informative view of variant differences.
-
 \subsection{Feature Search Extension}
@@
 \end{itemize}
+
 \subsection*{4.3.3 Feature Search Extension – Technical Description}
@@
 \section {References}
-\begin{thebibliography}{9}
-\bibitem{ucsc_trix}
-University of California, Santa Cruz. (n.d.). Trix Index Files. UCSC Genome Browser.
-\end{thebibliography}
+
 \end{document}

### 2025-05-05 UserManual/UserMan.tex
Adding to user manual script

@@
 \section{Use Cases - Video transcript}
+1a. TB Homepage & main page
+On opening of TersectBrowser+, we are met with the introductory page displaying the loaded datasets.
+The user can click on one of the available datasets, which will bring them to the main TersectBrowser+ exploratory page.
+The main body of the page is composed of a heatmap, comparing all accessions within the dataset against the selected reference accession.
+Darker bins represent areas with greater variation between the accession and the reference, whilst lighter bins represent more similar sequences.
+Vertical red bars signify gaps in the reference genome.
+Above the heatmap is a scale bar indicating chromosome position.
+
+1b. TB Main page header and buttons
+The menu bar above the heatmap contains navigational controls. Newly added in the TersectBrowser+ release is the `binoculars' button, the `Open Variant Search' button, and the `refresh' button.
+
+2a. Genome browser main window
+To view gene models corresponding to chromosome position, the user can click on the `binoculars' button, which will open the genome browser panel above the heatmap, showing gene models for the corresponding chromosomal positions.
+The panel has a left-margin offset to ensure the chromosomal scale bar between the genome browser panel and heatmap are aligned.
+In this view, the genome browser panel and heatmap are fully synced, such that any changes made to the horizontal scroll along the x-axis, zoom level,bin size, selected interval, or selected chromosome, will be immediately reflected in both the heatmap and genome browser panel.
+
+The initial view will show an error message, prompting the user to zoom in on decrease the bin size, as the default view of TersectBrowser+ is too zoomed out.
+The genome browser panel has a set height that can be changed by clicking and dragging the bottom of the panel.
+The user can also scroll vertically along the y-axis to view all gene models without changing the view of the TersectBrowser+ heatmap.
+
+2b. Genome Browser popup window
+If the user is interested in a specific accession or a specific bin, they can click on the specific accession or bin, which will open a popup window displaying a `View in browser' button.
+Clicking on this button will open a large popup window that is detached from the background TersectBrowser+, allowing the user to change the view by zooming or scrolling in the popup window without changing the view in TersectBrowser+.
+If the user has opened the genome browser popup window from accession name, the window will be fully zoomed out and show the entire chromosome.
+If the user has opened the genome browser popup window from a specific bin, the window will be zoomed in to show that particular interval.
 

### 2025-05-05 TechDoc/TechDoc.tex
Adjust architecture section

@@
 
-The frontend is a browser-based interface developed in Angular and React, designed for interactive data visualization, including distance plots, accession selectors, and genome browser integration.
+The frontend is a browser-based interface developed in Angular JS designed for interactive data visualization, including distance plots, accession selectors.
+
  The backend, built in Node.js and Express, provides API endpoints, data models, and utility scripts to serve genomic data and computation results from a MongoDB database.
+
  A set of common modules supports both frontend and backend logic with shared resources.
-Additionally, the architecture supports pluggable extensions, such as the genome-browser, which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons.
+Additionally, the architecture supports pluggable extensions, such as the genome-browser which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons.
+
  Supporting data files, environment configurations, and Python scripting for preprocessing or bioinformatics workflows are organized in separate folders for data loading and virtual environments, enabling efficient pipeline integration and deployment.
+
+\subsection{Embedding The Extensions }
 
@@
 
-\textit{Tersect Browser+} is primarily built using Angular 8, preserving the architecture of the original application.
-To enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22.
-This modular, "bolt-on" extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend and backend technologies.
+\textit{Tersect Browser+} is originally built using Angular 8, to enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22.
+
+New extensions are converted to web components and bundled with rollup [TODO: Reference] to a single min.js bundle, the target bundle is then transpiled to a version compatible with older browsers with babel [TODO: Reference].
+The final output is deployed on NPM.
+
+The genomebrowser relies on \texttt{Jbrowse} [TODO: Reference] extensively.
+Since Jbrowse is written in \texttt{React} [TODO: Reference], it is embedded in the new angular app via the \texttt{ReactComponentDirective} provided by the \texttt{@ngeenx/ngx-react} package on NPM.
+
+This modular,” bolt-on” extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend technologies.
+
 Extensions are developed independently and published to NPM, making them easily linkable through project dependencies.
 This hybrid approach ensures forward compatibility and scalability, while maintaining stability in the core application.
-Future versions of \textit{Tersect Browser+} may explore fully migrating the main application to a more unified and modern framework.
+Future versions of Tersect Browser+ may explore fully migrating the main application to a more unified and modern framework.
 
@@
 The relevance of the gene models in the GFF file depend on what the user provides in the dataset, and how the GFF file was created.
-
+\clearpage
 \subsection{Feature Search Extension}
@@
 
-\subsection*{4.3.3 Feature Search Extension – Technical Description}
+\subsection{Feature Search Extension - Technical Description}
 
-\subsubsection*{4.3.3.1 Creating Search Indexes Per Chromosome}
+\subsubsection{Creating Search Indexes Per Chromosome}
 
@@
 The final VCF was then compressed using \texttt{bgzip} and indexed with \texttt{tabix}.
-Additionally, Trix index files \citep{ucsc_trix} were generated for high impact variants on each chromosome to support rapid gene name search suggestions.
+Additionally, Trix index files [TODO:reference] were generated for high impact variants on each chromosome to support rapid gene name search suggestions.
 
-\subsubsection*{4.3.3.2 Handling Search Request}
+\subsubsection{Handling Search Request}
 
@@
 
-\subsubsection*{4.3.3.3 Handling Searching by Interval}
+\subsubsection{Handling Searching by Interval}
 
@@
 
-\subsubsection*{4.3.3.4 Handling Searching by Gene Name}
+\subsubsection{Handling Searching by Gene Name}
 
@@
 
-\subsubsection*{4.3.3.5 Linking with the Frontend}
+\subsubsection{Linking with the Frontend}
 
@@
 
-\subsubsection*{4.3.3.6 Other Ways to Trigger Variant Search}
+\subsubsection{Other Ways to Trigger Variant Search}
 
@@
 
-\subsubsection*{4.3.3.7 Highlighting Bins}
+\subsubsection{Highlighting Bins}
 
@@
 
-\subsubsection*{VCF Mechanism Summary}
+\subsubsection{VCF Mechanism Summary}
 

### 2025-05-05 UserManual/UserMan.tex
Added feature search sections

@@
 If the user has opened the genome browser popup window from a specific bin, the window will be zoomed in to show that particular interval.
+
+3. Feature Search
+The Feature Search functionality enables the user to search for specific variants (High/Moderate/Low) present within all accessions across a specified interval or within a specific gene.
+Clicking on the `Open Variant Search' button opens a popup window where the user can specify an interval to search, either by using the sliding bar or by manually typing in numbers, and can specify an impact level from the dropdown menu.
+Alternatively to searching by interval, the user can search by gene ID.
+The gene search bar has an autocomplete functionality and will suggest all possible gene IDs alphabetically.
+Searching by interval and gene ID are mutually exclusive, and when the user starts to type a gene ID the interval functionality becomes greyed out.
+To initiate the variant search, the user clicks on the `Search for variants' button, which highlights bins in the canvas representing the positions and accessions where variants of the specified impact level are present.
+
+Additionally, the user can select a specific bin or click and drag the chromosome scale bar to select an interval, which will bring up a popup menu with a `Search for variants' button.
+Clicking on this button will automatically open the `Variant Search' popup window with the interval preselected, and will automatically initiate the variant search.
+Again, bins in the heatmap will be highlighted red if they contain a variant of the specified impact level.
+
+4. Barcode generation
+
+5. Licensing and contact
 

### 2025-05-05 UserManual/UserMan.tex
Added barcode generation transcript

@@
 4. Barcode generation
+Once the user has identified an accession and a bin of interest, TersectBrowser+ offers the capability to generate accession-specific barcodes for that interval.
+Clicking on a specific bin opens a popup menu with a `Create barcode' button.
+Clicking this button opens a larger Barcode Generator popup window, where the user can specify barcode size and the maximum number of variants per barcode, which are populated with the default values on 150 and 1 respectively.
+Clicking on the `Generate' button will create the barcodes, and a `Download' button will appear.
+Clicking the `Download' button automatically downloads a .tsv file containing the generated barcodes and their associated statistics.
 

### 2025-05-05 UserManual/UserMan.tex
Added licensing and contact

@@
 
-5. Licensing and contact
+5. Loading datasets
+
+6. Licensing and contact
+TersectBrowser+ is an application developed by Cranfield University.
+It is developed under the MIT license and incorporates components from JBrowse, which is released under the Apache License 2.0.
+The source code is freely available at [Github link].
+The accompanying README contains additional information to setup and run TersectBrowser+.
+For future correspondance please contact Dr Tom Kurowski at: t.kurowski@cranfield.ac.uk.
 

### 2025-05-05 TechDoc/TechDoc.tex

@@
 
-The extensions were designed and added by the named authors under the supervision of of Dr Tomasz Kurowski between March and May 2025.
+The extensions were designed and added by the named authors under the supervision of Dr Tomasz Kurowski between March and May 2025.
 
@@
 
-\subsection{Embedding The Extensions }
+\subsection{Embedding The Extensions}
 
@@
 
-The genomebrowser relies on \texttt{Jbrowse} [TODO: Reference] extensively.
+The genome browser relies on \texttt{Jbrowse} [TODO: Reference] extensively.
 Since Jbrowse is written in \texttt{React} [TODO: Reference], it is embedded in the new angular app via the \texttt{ReactComponentDirective} provided by the \texttt{@ngeenx/ngx-react} package on NPM.
@@
     \item (3) identify areas where high-impact variants are visible; 
-    \item (4) be able to genereate DNA barcodes for the varioius cultivars and species in the dataset; and,  
+    \item (4) be able to generate DNA barcodes for the various cultivars and species in the dataset; and,  
     \item (5) to be able to load additional datasets, including for other crops. 
@@
     \item nvm (versions 16 and 22 specifically, as these will be deployed separately)
-    \item npm registery \url{https://registry.npmjs.org/}
+    \item npm registry \url{https://registry.npmjs.org/}
     \item Angular CLI v1.7.1
@@
 This is where the server will look to find information on tracks when generating the Variant Browser panel.
-            \item This script checks for the install of modules such as SAMtools, BCFtools, and HTSlib, and will attempt to download these if not found.
+            \item This script checks for the installation of modules such as SAMtools, BCFtools, and HTSlib, and will attempt to download these if not found.
 If a multi-sample VCF file is provided, it will split this file into individual sample level VCFs before proceeding.
-It will make sure these files are unzipped before using TERSECT CLI to build a tersect index for the whole dataset (it sources a previously created .tersect virtual environment to run this).
+It will make sure these files are unzipped before using TERSECT CLI to build a tersect index for the whole dataset (it sources a previously created tersect virtual environment to run this).
 After recompressing and indexing the fasta and VCF files, it will edit the \verb+add_example_dataset.sh+ script to point to the correct index file, and begin dataset addition to TersectBrowser+.
@@
         \verb+npm start+
-    \item Change to the ersect-browser/extension/genome-browser/ directory and run:\\
+    \item Change to the Tersect-browser/extension/genome-browser/ directory and run:\\
         \verb+nvm use 22+\\
@@
 
-\textit{Tersect Browser+} comprises Tersect CLI, Tersect Browser and four extensions that meet the \textit{Teresect Browser+} use case in Section 3. These were:
+\textit{Tersect Browser+} comprises Tersect CLI, Tersect Browser and four extensions that meet the \textit{Tersect Browser+} use case in Section 3. These were:
 
@@
 
-On reviewing node-compatible browser tools such as CRAMER, JBrowse was selected as the prefered solution, due to its high customisability, up-to-date features and active technical support.
+On reviewing node-compatible browser tools such as CRAMER, JBrowse was selected as the preferred solution, due to its high customisability, up-to-date features and active technical support.
 The initial build of JBrowse into \textit{TersectBrowser+}, focused on retaining features most relevant to the user aims.
-This includeed bringing the styling of the panel in line with the rest of the browser page.
+This included bringing the styling of the panel in line with the rest of the browser page.
 The extension also involved generating a popup window of the same JBrowse component to be initialised when the user identified a specific accession or bin region of interest.
@@
 The main window is located at the top of Tersect Browser, below the settings bar but above the heatmap.
-Main window shows only sequence tracks and linear genome scale.
-Interactive features such as zoom anc chromosome selection have been removed from the main window.
+Interactive features such as zoom and chromosome selection have been removed from the main window.
 The container window has a dynamic offset, which changes according to tree view and zoom level to ensure the chromosome scales between TersectBrowser and JBrowse line up.
 By default, the main window shows only the gene models track.
-
-The tracks include: 
-\begin{itemize}
-    \item Reference sequence 
-    \item Variant tracks for each loaded accession
-    \item Gene models FeatureTrack (\nameref{sec:Models})
-\end{itemize}
 
@@
 
-\subsubsection{Jbrowser TersectBrowser Syncing}
+\subsubsection{Jbrowse TersectBrowser Syncing}
 
@@
 \paragraph{View State} 
-The \verb +JBrowseLinearGenomeView+ component is imported from @jbrowse/react-linear-genome-view and is wrapped using React.
+The \verb +JBrowseLinearGenomeView+ component is imported from \verb*|@jbrowse/react-linear-genome-view| and is wrapped using React.
 The configurations for the assembly, assembly, and variant tracks, and tracks, are imported from the respective assembly.ts and tracks.ts configuration files.
@@
 
-\subsection{Feature Search Extension - Technical Description}
+\subsubsection{Technical Components}
 
@@
 
+The preparation process is automated with the \verb*|{build_annotate_per_chrom.sh}| script (Supplementary File 1). Its example usage is;
+
+\begin{verbatim}
+    build_annotate_per_chrom.sh tracks.json index.tsi genes.gff3.gz [outdir]
+\end{verbatim}
+
+Where \texttt{tracks.json} is the JBrowse tracks configuration file (TODO: Reference), \texttt{index.tsi} is the Tersect index file, \texttt{genes.gff3.gz} is the compressed GFF3 gene annotation file, and \texttt{[outdir]} is the output directory for the generated per-chromosome VCF and annotation files.
+
 The final VCF was then compressed using \texttt{bgzip} and indexed with \texttt{tabix}.
@@
 
-\subsubsection{Handling Search Request}
+The Trix file generation is automated with the \verb*|(tersect_trix_builder.sh)| script (Supplementary File 2).
 
-An API endpoint was added to the \texttt{tersect-router} backend to support search functionality. The structure of the endpoint is as follows:
+\subsubsection{Handling A Search Request}
+
+An API endpoint was added to the \texttt{tersect-router} file on the backend to support search functionality. The structure of the endpoint is as follows:
 
@@
 
-\subsubsection{Handling Searching by Interval}
+\subsubsection{Searching by Interval}
 
@@
 
-To associate variants with accession names, a function uses \texttt{tersect} to determine which accessions have each specific variant.
-A function to get the optimal concurrency (\texttt(getOptimalConcurrency)) was introduced which detected the number of cores present on the server, subtracting 1 from the total and using the rest.
+To associate variants with accession names, tersectForKey function uses \texttt{tersect} to determine which accessions have each specific variant.
+A function to get the optimal concurrency \texttt(getOptimalConcurrency) was introduced which detected the number of cores present on the server, subtracting 1 from the total and using the rest.
 To prevent server overload from concurrent \texttt{tersect} calls, the \texttt{p-limit} library is used to control concurrency based on available CPU cores.
@@
 
-\subsubsection{Handling Searching by Gene Name}
+\subsubsection{Searching by Gene Name}
 
@@
 
-Variant searches can also be triggered from:
+Variant search can also be triggered from:
 \begin{itemize}
@@
 
-These triggers emit a custom event \texttt{tersect-search-variants}, which sets the interval and opens the modal for user interaction.
+These triggers emit a custom event \texttt{tersect-search-variants}, which is being listened for on the react component.
+This sets the interval in the search dialog and opens the modal for user interaction.
 
@@
 The top-level function \texttt{callHighlightBins(searchedAccessions)} is defined in \texttt{tersect-browser.component.ts}, where it retrieves interval, bin size, and accession info, and calls \texttt{highlightBins()}.
-
-\subsubsection{VCF Mechanism Summary}
-
-The overall mechanism for variant search begins with API queries, uses either \texttt{bcftools} or Trix for data retrieval, links variants to accessions with \texttt{tersect}, and finally triggers the highlighting of bins in the Tersect UI.
-
-\paragraph{Other Triggers for Variant Search}
-Variant search also triggered via bin plot or scale plot selection using the \texttt{tersect-search-variants} custom event.
 
